---
title: "Curi Draft"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library(tidyverse)
library(MutationalPatterns)
library(NMF)
library(reshape2)
library(grid)
library(gridExtra)
library(ggpubr)
library(ggrepel)
library(broom)
library(emplik)
library(glmc)
library(ConsReg)
library(Hmisc)
library(yardstick)
library(reshape2)
```

## Preliminary functions

The following functions allow us to load and analyze mutational signatures

```{r}
# Loads a mutational signature matrix from a mutational signature txt file
load_old_mutational_matrix<- 
  function(filename,
          mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv") {
  mutational.signature <-  read_csv(filename)
  mut_mat <-  read_csv(mut.order)

  mut.join <-  mutational.signature %>%
    inner_join(mut_mat)
  
  order.tri <-  unique(mut.join$ext.context)

  mut.join <-  mut.join %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    select(ext.context,prob)
  
  mut.matrix <-  as.matrix(
        sapply(mut.join[,2:ncol(mut.join)], as.numeric))  

  rownames(mut.matrix) <-  mut.join$ext.context
 
  return (t(mut.matrix)[,1:96])
}

change_old_mutational_matrix<-
    function(sig.frame,
             mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv") {
        
  mutational.signature <-  sig.frame
  mut_mat <-  read_csv(mut.order)

  mut.join <-  mutational.signature %>%
    inner_join(mut_mat)
  
  order.tri <-  unique(mut.join$ext.context)

  mut.join <-  mut.join %>%
      mutate(ext.context =
                 factor (ext.context, levels = order.tri)) %>%
      select(sample,ext.context,prob) %>%
      pivot_wider(names_from = ext.context,
                  values_from = prob)
    
    mut.matrix <-  as.matrix(
        sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
    rownames(mut.matrix) = mut.join$sample

    return (t(mut.matrix))
}

load_cosmic_matrix<- function(mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv") {

  sp_url <- paste("https://cancer.sanger.ac.uk/cancergenome/assets/",
                "signatures_probabilities.txt", sep = "")
   
  cancer_signatures <-  read.table(sp_url, sep = "\t", header = TRUE)
  # Match the order of the mutation types to MutationalPatterns standard
  mut_mat <-  read_csv(mut.order)
  new_order <-  match(mut_mat$ext.context, cancer_signatures$Somatic.Mutation.Type) 
  # Reorder cancer signatures dataframe
  cancer_signatures <-  cancer_signatures[as.vector(new_order),]
  # Add trinucletiode changes names as row.names
  row.names(cancer_signatures) <-  cancer_signatures$Somatic.Mutation.Type
  # Keep only 96 contributions of the signatures in matrix
  cancer_signatures <- as.matrix(cancer_signatures[,4:33])
  return(cancer_signatures)
}

# Returns a frame with columns Signature, Sample, Contribution with the absolute
# contribution of each signature for each sample
calc_abs_contribution <- function(nmf.res,sig.names) {

  total_signatures <-  colSums(nmf.res$signatures) 
  abs_contribution <-  nmf.res$contribution * total_signatures
  rownames(abs_contribution) <- sig.names

  m_contribution <-  melt(abs_contribution)
  colnames(m_contribution) = c("signature", "sample", "contribution")

  return (m_contribution)
}

# Returns a frames with the cosine similarity (fit) between the original profile and the
# reconstructed profile
calc_fit_contribution <- function(mut.matrix,nmf.res) {

  # calculate all pairwise cosine similarities
  cos_sim_ori_rec <- cos_sim_matrix(mut.matrix, nmf.res$reconstructed)
  # extract cosine similarities per sample between original and reconstructed
  fit.table<- as.data.frame(diag(cos_sim_ori_rec))
  fit.table$Sample <-  rownames(fit.table)
  colnames(fit.table) = c("fit","sample")

  return(fit.table)
}

get_top_signature <-function(mut.table, signatures.table,
                             signature,top.n=8,
                             signatures=
                                 c("PolE","Aging","MMR","Germline")){

    # Construct table with only contribution from specific signature
    sig.table <-  mut.table %>%
        gather(signatures,key="sign.name",value="contribution") %>%
        filter(sign.name==signature) %>%
        arrange(desc(contribution))

    # Pick top n samples for that signature
    top.table <- 
        sig.table %>% arrange(desc(contribution)) %>%
        head(top.n)

    # Pick signatures for those top samples and plot them
    mut.selected <- top.table %>%
        select(sample) %>%
        inner_join (signatures.table)

    top.signatures <-  change_old_mutational_matrix(mut.selected)
    colnames(top.signatures) <-  paste0("P",1:top.n,"")

    plot.top <-  plot_96_profile(top.signatures)

    return (plot.top)
}
```

## Data acquisition and cleaning

We load the mutational signatures file for all of the samples and consider only 
samples that have high sequencing coverage (over 5Mb at 20x)
```{r}
# Load mutational signature data for all samples
signatures <- read_csv(
    "mutational_sig_oec-main/supplied_data/mutational.sig.csv")

# Load sequencing/cohort info for samples, filter with at
# at least 5 million Mb covered
sample.info <-
    read_csv("mutational_sig_oec-main/supplied_data/sample.info.csv") %>%
    filter (area.cov > 5e6)
```

## Linear regression model
This code generates the simple Linear regression model for profile p3 based on Signature 10 and Signature6 or Signature1.
```{r}
#Load the P3, Sig10,6,5 data
data <- read_csv("Data/P3_sig.csv")

model1 <- lm(p3 ~ 0 + sig10 + sig6, data = data)
summary(model1)

model2 <- lm(p3 ~ 0 + sig10 + sig1, data = data)
summary(model2)

#plot1
ggplot(data, aes(sig10, p3)) +
  geom_point() +
  geom_smooth(method = "lm")

data1 <- data %>%
  pivot_longer(
    cols = c(`sig10`, `sig6`, `sig1`),
    names_to = "signatures", 
    values_to = "prob"
  )

#plot2
ggplot(data1, aes(x = prob, y = p3, color = signatures)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    y = "Mutation prob of P3",
    x = "Mutation prob of Sig",
    title = "Contribution of Signatures to the Profile P3"
  ) +
  theme(
        axis.text.x = element_text(color = "grey20", size = 17, angle = 0, hjust = 1, vjust = 0, face = "plain"), axis.text.y = element_text(color = "grey20", size = 17, angle = 0, hjust = 1, vjust = 0, face = "plain"), axis.title = element_text(size = 17), plot.title = element_text(size=22), legend.text=element_text(size=17), legend.title =element_text(size=17)) 
```


## Function for generating the sample mutation profiles
Load the Mutation signatures order
```{r}
mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv"
mutational.signature <-  mut.selected
mut_mat <-  read_csv(mut.order)

mut.join <-  mutational.signature %>%
  inner_join(mut_mat)
order.tri <-  unique(mut.join$ext.context)
mut_order <- as_tibble(order.tri)
```


This function iss for generating the sample profile with multiple signatures.
```{r, message=FALSE, warning=FALSE}
#sigs = the array of the signatures that we want to put 
#nums = the number of mutations that we want to generate for each signature

generate_sample_prof <- function(sigs, nums){
  
  cosmic.signatures <- get_known_signatures()
  sig <- cosmic.signatures[,sigs]
  
  #load the mutation order
  mut.selected <- sample.info %>%
    select (sample) %>%
    inner_join (signatures)
  
  mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv"
  mutational.signature <-  mut.selected
  mut_mat <-  read_csv(mut.order)
  
  mut.join <-  mutational.signature %>%
    inner_join(mut_mat)
  order.tri <-  unique(mut.join$ext.context)
  
  if(length(sigs) == 1){
    names(sig) <- order.tri
    sig <- as.matrix(sig)
  } else{
    rownames(sig) <- order.tri
    mut_order <- as_tibble(order.tri)
  }
  
  df_a <- c()
  #generate from the first signature to the last signatures and merge with the first signature
  for (i in 1:length(sigs)){
    b <- sample(rownames(sig), size = nums[i], replace = TRUE, prob = sig[,i])
    df_b <- as_tibble(b)
    df_b <- df_b %>%
      count(value) %>%
      mutate(
        value = factor(value, levels = order.tri),
      ) %>%
      merge(mut_order, all = TRUE)
    
    df_b[is.na(df_b)] <- 0     # Replace NA with 0
    df_b <- df_b %>%
      select(-value)
    
    #rownames(df_b) <- order.tri
    df_b <- as.matrix(df_b)
    
    df_a <- cbind(df_a, df_b)
  }
  
  df_a <- rowSums(df_a)
  df_a <- as_tibble(df_a)
  colnames(df_a) <- c("mut")

  sample_profile <- df_a %>%
    mutate(
      prob = mut/sum(mut)
    )
  sample_profile <- as.matrix(sample_profile)
  rownames(sample_profile) <- order.tri
  
  sample_profile
}
```

This function is plotting the sample mutational profile based on the selected signatures and the number of mutations that we want to generate for each signature.
```{r, message=FALSE, warning=FALSE}
#sigs = the array of the signatures that we want to put 
#nums = the number of mutations that we want to generate for each signature

plot_sample_prof <- function(sample_profile){
  
  sample_profile <- sample_profile[,2]
  sample_profile <- as.matrix(sample_profile)
  colnames(sample_profile) <- c("prob")
  
  #plot the matrix or tibble
  plot_96_profile(sample_profile)
}
```

## Linear Regression Model with constraints
This function generates the linear combination of the selected multiple signatures with positive constraints for the generated sample mutational profile.

```{r}
estimate_coeff <- function(sample_profile, sigs, tissue_type, scale, quantiles, mb_size){
  
  #parameters testing
  #sample_profile <- generate_sample_prof(s_list, m_list)
  total_mut <- unname(colSums(sample_profile)[1])
  
  sample_profile <- as.matrix(sample_profile[,2])
  colnames(sample_profile) <- c("prob")
  
  # sigs <- s_list
  # tissue_type <- "COAD"
  # scale <- "exome"
  # quantiles <- c("5%","95%")
  # mb_size <- 50

  #pulling constraints
  constraints_ds <- read_csv("Data/constraints_ds.csv")
  constr1 <- constraints_ds %>%
    filter(Tissue == tissue_type, Signature %in% sigs, Scale == scale, Quantile %in% quantiles) 
  constrs <- c(constr1$Value)
  
  #making dataset for the model
  cosmic.signatures <- get_known_signatures()
  sig <- cosmic.signatures[,sigs]
  
  df <- cbind(sample_profile,sig)
  if(length(sigs) == 1){
    colnames(df) <- c("prob", sigs)
  }
  df <- as_tibble(df)
  
  if(length(sigs) == 1){
    constraints <- str_c(constrs[1]*(mb_size/total_mut), ' < ', sigs[1], ', ', sigs[1], ' < ', constrs[2]*(mb_size/total_mut))
  } else{
        constraints <- str_c(constrs[1]*(mb_size/total_mut), ' < ', sigs[1], ', ', sigs[1], ' < ', constrs[2]*(mb_size/total_mut))
    for (i in 2:length(sigs)){
      constraint <- str_c(constrs[2*i-1]*(mb_size/total_mut), ' < ', sigs[i], ', ', sigs[i], ' < ', constrs[2*i]*(mb_size/total_mut))
      constraints <- str_c(constraints, ', ', constraint)
    }
  }
  
  model<- ConsReg(formula = prob ~ 0 + ., data = df,
            family = 'gaussian',
            constraints = constraints,
            LOWER = 0, UPPER = 1.1
            )
  
  model$coefficients
}
```

This function is our final method named **Dissolvo** and prevents the model from having negative coefficients.
//
```{r}
sample_profile <- generate_sample_prof(c("SBS1","SBS10b","SBS6"),c(200,300,500))
test1 <- estimate_coeff(sample_profile, sigs = c("SBS1","SBS10b","SBS6"), tissue_type = "COAD", scale = "exome", quantiles = c("40%", "50%"), mb_size = 50)
test1

dissolvo <- function(sample_profile, sigs, tissue_type, scale, quantiles, mb_size){
  neg_count <- 1
  while(neg_count != 0){
    sample_prof <- generate_sample_prof(s_list, m_list)
    model <- estimate_coeff(sample_prof, sigs = s_list, tissue_type, scale, quantiles, mb_size)
    coeff_list <- unname(model)
    neg_count <- sum(coeff_list < 0)
  }
  model
}
```


Method testing code
```{r}
#1 sig
sample_profile <- generate_sample_prof(c("SBS10a"),c(7000))
plot_sample_prof(sample_profile)
estimate_coeff(sample_profile, sigs = c("SBS10a"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)
dissolvo(c("SBS10a"),c(7000), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#2 sigs
sample_profile <- generate_sample_prof(c("SBS10a","SBS6"),c(3000,7000))
plot_sample_prof(sample_profile)
estimate_coeff(sample_profile, sigs = c("SBS10a","SBS6"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)
dissolvo(c("SBS10a","SBS6"), c(3000,7000), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#3 sigs
sample_profile <- generate_sample_prof(c("SBS1","SBS10b","SBS6"),c(200,300,500))
plot_sample_prof(sample_profile)
estimate_coeff(sample_profile, sigs = c("SBS1","SBS10b","SBS6"), tissue_type = "COAD", scale = "exome", quantiles = c("40%", "50%"), mb_size = 50)
dissolvo(c("SBS1","SBS10b","SBS6"), c(200,300,500), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#4 sigs
sample_profile <- generate_sample_prof(c("SBS10a","SBS10b","SBS1","SBS6"),c(300,100,200,400))
plot_sample_prof(sample_profile)
estimate_coeff(sample_profile, sigs = c("SBS10a","SBS10b","SBS1","SBS6"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)
dissolvo(c("SBS10a","SBS10b","SBS1","SBS6"), c(300,100,200,400), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#5 sigs
sample_profile <- generate_sample_prof(c("SBS10a","SBS10b","SBS1","SBS6", "SBS2"),c(200,300,100,300,100))
plot_sample_prof(sample_profile)
estimate_coeff(sample_profile, sigs = c("SBS10a","SBS10b","SBS1","SBS6", "SBS2"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)
dissolvo(c("SBS10a","SBS10b","SBS1","SBS6", "SBS2"),c(200,300,100,300,100), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)
```



## Evaluation of Dissolvo
```{r}
COSMICsample_ds <- read_csv("Data/COSMICsample_ds.csv")
cosmic.signatures <- get_known_signatures()
tissue_list <- unique(COSMICsample_ds$Cancer.Types)
```

These are the signatures that the constraints are unknown in the COSMIC website.
```{r}
COSMICsample <- COSMICsample_ds %>%
  select(6:68) 
signature_list <- colnames(COSMICsample)
signature_list <- str_remove(signature_list, ".Mutations")
cosmic_signature_list <- colnames(cosmic.signatures)

known_sig <- signature_list[signature_list %in% cosmic_signature_list]
```
```{r}
coloE_samples <- COSMICsample_ds %>%
  filter(Cancer.Types == "ColoRect-AdenoCA") %>%
  filter(Scale == "exome")

coloE_samples_withnames <- coloE_samples %>%
  select(4,6:68)
coloE_samples <- coloE_samples %>%
  select(6:68) 
coloE_size <- dim(coloE_samples)[1]

sample_names_list <- coloE_samples_withnames %>%
    select(1)
sample_names_list <- as.matrix(sample_names_list)
```

This chunk is making the final dataframe for the evaluation of the method.

//fixing
```{r, eval=FALSE}
actual_list <- c()
expected_list <- c()
fitsig_list <- c()
sig_list <- c()
sample_list <- c()

for (i in 1:coloE_size){
  
  mut <- as.numeric(coloE_samples[i,])
  nonzero_sig <- (which(mut %nin% c(0, NA)))
  list <- colnames(coloE_samples)
  s_list <- list[nonzero_sig]
  s_list <- str_remove(s_list, ".Mutations")
  m_list <- mut[mut %nin% c(0, NA)]
  
  bool <- s_list %in% known_sig
  s_list <- s_list[bool]
  m_list <- m_list[bool]
  
  if(length(m_list) == 1){
    next
  }
  
  prop.test <- prop.table(m_list)
  
  sample_prof <- generate_sample_prof(s_list, m_list)
  method_test <- estimate_coeff(sample_prof, sigs = s_list, tissue_type = "COAD", scale = "exome", quantiles = c("5%","95%"), mb_size = 50)
  fit_test <- fit_to_signatures(sample_prof, cosmic.signatures[,s_list])
  fit_test <- fit_test$contribution[,2]
  
  sample_names <- c()
  for(j in 1:length(s_list)){
    sample_name <- sample_names_list[i]
    sample_names <- append(sample_names, sample_name)
  }
  
  sample_list <- append(sample_list, sample_names)
  actual_list <- append(actual_list, prop.test)
  expected_list <- append(expected_list, method_test)
  fitsig_list <- append(fitsig_list, fit_test)
  sig_list <- append(sig_list, s_list)
}

eval_df <- cbind(sample_list, sig_list, expected_list, fitsig_list, actual_list)
eval_df <- as_tibble(eval_df)
eval_df <- eval_df %>%
  mutate(
    expected_list = as.numeric(expected_list),
    fitsig_list = as.numeric(fitsig_list),
    actual_list = as.numeric(actual_list)
  )
``` 

```{r, eval=FALSE}
actual_list <- c()
expected_list <- c()
fitsig_list <- c()
sig_list <- c()
sample_list <- c()

for (i in 1:coloE_size){
  
  mut <- as.numeric(coloE_samples[i,])
  nonzero_sig <- (which(mut %nin% c(0, NA)))
  list <- colnames(coloE_samples)
  s_list <- list[nonzero_sig]
  s_list <- str_remove(s_list, ".Mutations")
  m_list <- mut[mut %nin% c(0, NA)]
  
  bool <- s_list %in% known_sig
  s_list <- s_list[bool]
  m_list <- m_list[bool]
  
  if(length(m_list) == 1){
    next
  }
  
  prop.test <- prop.table(m_list)
  
  neg_count <- 1
  while(neg_count != 0){
    sample_prof <- generate_sample_prof(s_list, m_list)
    model <- estimate_coeff(sample_prof, sigs = s_list, tissue_type = "OV", scale = "exome", quantiles = c("5%","95%"), mb_size = 50)
    coeff_list <- unname(model)
    neg_count <- sum(coeff_list < 0)
  }
  method_test <- model

  fit_test <- fit_to_signatures(sample_prof, cosmic.signatures[,s_list])
  fit_test <- fit_test$contribution[,2]
  
  sample_names <- c()
  for(j in 1:length(s_list)){
    sample_name <- sample_names_list[i]
    sample_names <- append(sample_names, sample_name)
  }
  
  sample_list <- append(sample_list, sample_names)
  actual_list <- append(actual_list, prop.test)
  expected_list <- append(expected_list, method_test)
  fitsig_list <- append(fitsig_list, fit_test)
  sig_list <- append(sig_list, s_list)
}

eval_df <- cbind(sample_list, sig_list, expected_list, fitsig_list, actual_list)
eval_df <- as_tibble(eval_df)
eval_df <- eval_df %>%
  mutate(
    expected_list = as.numeric(expected_list),
    fitsig_list = as.numeric(fitsig_list),
    actual_list = as.numeric(actual_list)
  )
``` 
```{r, eval=FALSE}
write.csv(eval_df, "Data/Ovarian_ds.csv", row.names = FALSE)
```

//
```{r}
eval_df %>%
  mutate(
    diff1 = abs(expected_list - actual_list),
    diff2 = abs(fitsig_list - actual_list)) %>%
  filter(diff1 < diff2) %>%
  filter(diff2 - diff1 > 0.01)

eval_df %>%
  mutate(
    diff1 = abs(expected_list - actual_list),
    diff2 = abs(fitsig_list - actual_list)) %>%
  filter(diff1 > diff2) %>%
  filter(diff1 - diff2 > 0.01)
```


Plot the Expected vs. Actual
```{r}
eval_df <- read_csv("Data/evaluation_ds.csv")

ggplot(data = eval_df, aes(x = expected_list, y = actual_list))+
   geom_point(alpha=0.3) +
   geom_smooth(method = "lm") +
   labs(x = "Expected Value", y = "Actual Value")

cor(eval_df$expected_list, eval_df$actual_list)
cor(eval_df$fitsig_list, eval_df$actual_list)

sig_levels <- c("SBS1", "SBS5", "SBS15", "SBS40", "SBS44")

sig_levels <- c("SBS1", "SBS5", "SBS15", "SBS40", "SBS44")
eval_df2 <- eval_df %>%
  #mutate(sig_list = factor(sig_list, sig_levels)) %>%
  group_by(sig_list) %>%
  summarise(n = n()) %>%
  filter(n > 65 & !is.na(sig_list))

ggplot(data = eval_df, aes(x = expected_list, y = actual_list))+
   geom_point(alpha=0.3) +
   geom_smooth(method = "lm") +
   facet_wrap(~sig_list) +
   labs(x = "Expected Value", y = "Actual Value")
```

order
sort out
```{r}
eval_df <- eval_df %>%
  mutate(diff = abs(actual_list - expected_list)) %>%
  arrange(desc(diff)) 

histogram(eval_df$diff)
```

Plot the Expected vs. fit_to_sig 
```{r}
ggplot(data = eval_df, aes(x = expected_list, y = fitsig_list))+
   geom_point(alpha=0.3) +
   geom_smooth(method = "lm") +
   labs(x = "Expected Value", y = "fit_to_sig Value")

ggplot(data = eval_df, aes(x = expected_list, y = fitsig_list))+
   geom_point(alpha=0.3) +
   geom_smooth(method = "lm") +
   facet_wrap(~sig_list) +
   labs(x = "Expected Value", y = "fit_to_sig Value")

cor(eval_df$expected_list, eval_df$fitsig_list)
```

Plot all together
```{r}
eval_df_plot <- eval_df %>%
  pivot_longer(
    cols = c(`expected_list`, `fitsig_list`),
    names_to = "method", 
    values_to = "coeff", 
    values_drop_na = TRUE
  )

ggplot(eval_df_plot, aes(actual_list, coeff, colour = method)) +
   geom_point(alpha=0.3) +
   geom_smooth(method = "lm") +
   labs(x = "Truth")

ggplot(eval_df_plot, aes(actual_list, coeff, colour = method)) +
   geom_point(alpha=0.3) +
   geom_smooth(method = "lm") +
   facet_wrap(~sig_list) +
   labs(x = "Truth")

histogram(eval_df$fitsig_list-eval_df$actual_list)
histogram(eval_df$expected_list-eval_df$actual_list)
mosaic::fav_stats(eval_df$fitsig_list-eval_df$actual_list)
mosaic::fav_stats(eval_df$expected_list-eval_df$actual_list)
```

histogram for all cancer types

Getting MSE between Expected vs. fit_to_signature vs. Actual
```{r, eval=FALSE}
mse_list_EF <- c()
mse_list_EA <- c()

for (i in 1:coloE_size){
  
  mut <- as.numeric(coloE_samples[i,])
  nonzero_sig <- (which(mut %nin% c(0, NA)))
  list <- colnames(coloE_samples)
  s_list <- list[nonzero_sig]
  s_list <- str_remove(s_list, ".Mutations")
  m_list <- mut[mut %nin% c(0, NA)]
  
  bool <- s_list %in% known_sig
  s_list <- s_list[bool]
  m_list <- m_list[bool]
  
  if(length(m_list) == 1){
    next
  }
  
  prop_test <- prop.table(m_list)
  
  neg_count <- 1
  while(neg_count != 0){
    sample_prof <- generate_sample_prof(s_list, m_list)
    model <- estimate_coeff(sample_prof, sigs = s_list, tissue_type = "COAD", scale = "exome", quantiles = c("5%","95%"), mb_size = 50)
    coeff_list <- unname(model)
    neg_count <- sum(coeff_list < 0)
  }
  method_test <- model
  
  fit_test <- fit_to_signatures(sample_prof, cosmic.signatures[,s_list])
  fit_test <- fit_test$contribution[,2]
  
  df <- cbind(fit_test, method_test)
  df <- as_tibble(df)
  mse <- rmse_vec(df$fit_test, df$method_test)^2
  mse_list_EF <- append(mse_list_EF, mse)

  df <- cbind(prop_test, method_test)
  df <- as_tibble(df)
  mse <- rmse_vec(df$prop_test, df$method_test)^2
  mse_list_EA <- append(mse_list_EA, mse)
}

coloE_sample_names <- unique(eval_df$sample_list)
mse_df <- cbind(coloE_sample_names, mse_list_EF, mse_list_EA)
mse_df <- as_tibble(mse_df)
mse_df <- mse_df %>%
  mutate(
    mse_list_EF = as.numeric(mse_list_EF),
    mse_list_EA = as.numeric(mse_list_EA)
  )
```
```{r, eval=FALSE}
write.csv(mse_df, "Data/mse_ds.csv", row.names = FALSE)
```

Plot of MSE between Expected vs. fit_to_signature vs. Actual
```{r}
mse_df <- read_csv("Data/mse_ds.csv")
```
```{r}
mean(mse_df$mse_list_EA)
mean(mse_df$mse_list_EF)

histogram(mse_df$mse_list_EA)
histogram(mse_df$mse_list_EF)

mse_df %>%
  ggplot(aes(x = seq(1, dim(mse_df)[1]), y = mse_list_EA)) +
    geom_point() +
    labs(x = "samples", y = "MSE", title = "MSE between Expected and Actual value") 

mse_df %>%
  ggplot(aes(x = seq(1, dim(mse_df)[1]), y = mse_list_EF)) +
    geom_point() +
    labs(x = "samples", y = "MSE", title = "MSE between Expected and fit_to_signature value") 
```




