---
title: "curi1"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library(tidyverse)
library(MutationalPatterns)
library(NMF)
library(reshape2)
library(grid)
library(gridExtra)
library(ggpubr)
library(ggrepel)
```

## Preliminary functions

The following functions allow us to load and analyze mutational signatures

```{r}
# Loads a mutational signature matrix from a mutational signature txt file
load_old_mutational_matrix<- 
  function(filename,
          mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv") {
  mutational.signature <-  read_csv(filename)
  mut_mat <-  read_csv(mut.order)

  mut.join <-  mutational.signature %>%
    inner_join(mut_mat)
  
  order.tri <-  unique(mut.join$ext.context)

  mut.join <-  mut.join %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    select(ext.context,prob)
  
  mut.matrix <-  as.matrix(
        sapply(mut.join[,2:ncol(mut.join)], as.numeric))  

  rownames(mut.matrix) <-  mut.join$ext.context
 
  return (t(mut.matrix)[,1:96])
}

change_old_mutational_matrix<-
    function(sig.frame,
             mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv") {
        
  mutational.signature <-  sig.frame
  mut_mat <-  read_csv(mut.order)

  mut.join <-  mutational.signature %>%
    inner_join(mut_mat)
  
  order.tri <-  unique(mut.join$ext.context)

  mut.join <-  mut.join %>%
      mutate(ext.context =
                 factor (ext.context, levels = order.tri)) %>%
      select(sample,ext.context,prob) %>%
      pivot_wider(names_from = ext.context,
                  values_from = prob)
    
    mut.matrix <-  as.matrix(
        sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
    rownames(mut.matrix) = mut.join$sample

    return (t(mut.matrix))
}

load_cosmic_matrix<- function(mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv") {

  sp_url <- paste("https://cancer.sanger.ac.uk/cancergenome/assets/",
                "signatures_probabilities.txt", sep = "")
   
  cancer_signatures <-  read.table(sp_url, sep = "\t", header = TRUE)
  # Match the order of the mutation types to MutationalPatterns standard
  mut_mat <-  read_csv(mut.order)
  new_order <-  match(mut_mat$ext.context, cancer_signatures$Somatic.Mutation.Type) 
  # Reorder cancer signatures dataframe
  cancer_signatures <-  cancer_signatures[as.vector(new_order),]
  # Add trinucletiode changes names as row.names
  row.names(cancer_signatures) <-  cancer_signatures$Somatic.Mutation.Type
  # Keep only 96 contributions of the signatures in matrix
  cancer_signatures <- as.matrix(cancer_signatures[,4:33])
  return(cancer_signatures)
}

# Returns a frame with columns Signature, Sample, Contribution with the absolute
# contribution of each signature for each sample
calc_abs_contribution <- function(nmf.res,sig.names) {

  total_signatures <-  colSums(nmf.res$signatures) 
  abs_contribution <-  nmf.res$contribution * total_signatures
  rownames(abs_contribution) <- sig.names

  m_contribution <-  melt(abs_contribution)
  colnames(m_contribution) = c("signature", "sample", "contribution")

  return (m_contribution)
}

# Returns a frames with the cosine similarity (fit) between the original profile and the
# reconstructed profile
calc_fit_contribution <- function(mut.matrix,nmf.res) {

  # calculate all pairwise cosine similarities
  cos_sim_ori_rec <- cos_sim_matrix(mut.matrix, nmf.res$reconstructed)
  # extract cosine similarities per sample between original and reconstructed
  fit.table<- as.data.frame(diag(cos_sim_ori_rec))
  fit.table$Sample <-  rownames(fit.table)
  colnames(fit.table) = c("fit","sample")

  return(fit.table)
}

get_top_signature <-function(mut.table, signatures.table,
                             signature,top.n=8,
                             signatures=
                                 c("PolE","Aging","MMR","Germline")){

    # Construct table with only contribution from specific signature
    sig.table <-  mut.table %>%
        gather(signatures,key="sign.name",value="contribution") %>%
        filter(sign.name==signature) %>%
        arrange(desc(contribution))

    # Pick top n samples for that signature
    top.table <- 
        sig.table %>% arrange(desc(contribution)) %>%
        head(top.n)

    # Pick signatures for those top samples and plot them
    mut.selected <- top.table %>%
        select(sample) %>%
        inner_join (signatures.table)

    top.signatures <-  change_old_mutational_matrix(mut.selected)
    colnames(top.signatures) <-  paste0("P",1:top.n,"")

    plot.top <-  plot_96_profile(top.signatures)

    return (plot.top)
}

```

## Data acquisition and cleaning

We load the mutational signatures file for all of the samples and consider only 
samples that have high sequencing coverage (over 5Mb at 20x)
```{r}
# Load mutational signature data for all samples
signatures <- read_csv(
    "mutational_sig_oec-main/supplied_data/mutational.sig.csv")

# Load sequencing/cohort info for samples, filter with at
# at least 5 million Mb covered
sample.info <-
    read_csv("mutational_sig_oec-main/supplied_data/sample.info.csv") %>%
    filter (area.cov > 5e6)


```

## Linear regression model
```{r}
#Load the P3, Sig10,6,5 data
data <- read_csv("mutational_sig_oec-main/supplied_data/P3_sig.csv")

model1 <- lm(p3 ~ 0 + sig10 + sig6, data = data)
summary(model1)

model2 <- lm(p3 ~ 0 + sig10 + sig1, data = data)
summary(model2)

#plot1
ggplot(data, aes(sig10, p3)) +
  geom_point() +
  geom_smooth(method = "lm")

data1 <- data %>%
  pivot_longer(
    cols = c(`sig10`, `sig6`, `sig1`),
    names_to = "signatures", 
    values_to = "prob"
  )

#plot2
ggplot(data1, aes(x = prob, y = p3, color = signatures)) +
  geom_point() +
  geom_smooth(method = "lm")
```

## Function testing code
```{r}
#load the Signatures 10, 6, 1
cosmic.signatures <- load_cosmic_matrix()
sig <- cosmic.signatures[,c(10,6,1)]
sig.labels <-  c("10","6","1")
colnames(sig) <- sig.labels

#load the mutation order
mut.selected <- sample.info %>%
  select (sample) %>%
  inner_join (signatures)

mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv"
mutational.signature <-  mut.selected
mut_mat <-  read_csv(mut.order)

mut.join <-  mutational.signature %>%
  inner_join(mut_mat)
order.tri <-  unique(mut.join$ext.context)
mut_order <- as_tibble(order.tri)

#generate random mutation samples of Signature 10
a <- sample(rownames(sig), size = 500, replace = TRUE, prob = as.vector(sig[,"10"]))
df_a <- as_tibble(a)
df_a <- df_a %>%
  count(value) %>%
  mutate(
    value = factor(value, levels = order.tri),
    prob = n/sum(n),
  ) %>%
  merge(mut_order, all = TRUE)

df_a[is.na(df_a)] <- 0     # Replace NA with 0
df_a <- df_a %>%
  mutate(
    prob = prob + 1e-04
  )

#generate random mutation samples of Signature 6
b <- sample(rownames(sig), size = 500, replace = TRUE, prob = as.vector(sig[,"6"]))
df_b <- as_tibble(b)
df_b <- df_b %>%
  count(value) %>%
  mutate(
    value = factor(value, levels = order.tri),
    prob = n/sum(n)
  ) %>%
  merge(mut_order, all = TRUE)

df_b[is.na(df_b)] <- 0     # Replace NA with 0
df_b <- df_b %>%
  mutate(
    prob = prob + 1e-04
  )

#join those two sample mutations
mut_df <- df_a %>%
  inner_join(df_b, by = "value") %>%
  mutate(
    prob = (prob.x + prob.y)/2
  ) %>%
  select(value, prob)

#define the sample profile created by two sample mutations
sample_profile <- mut_df %>%
  select(prob)
sample_profile <- as.matrix(sample_profile)
```

## Function for generating the sample mutation profiles
Load the Mutation signatures order
```{r}
mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv"
mutational.signature <-  mut.selected
mut_mat <-  read_csv(mut.order)

mut.join <-  mutational.signature %>%
  inner_join(mut_mat)
order.tri <-  unique(mut.join$ext.context)
mut_order <- as_tibble(order.tri)
```

Function for generating the sample mutation profiles 
```{r}
#A, B = Signatures.A, Signatures.B
#n = the number of mutations that we want to generate
generate_sample_prof <- function(A, B, n){

  cosmic.signatures <- load_cosmic_matrix()
  sig <- cosmic.signatures[,c(A, B)]
  sig.labels <-  c("A","B")
  colnames(sig) <- sig.labels
  
  #load the mutation order
  mut.selected <- sample.info %>%
    select (sample) %>%
    inner_join (signatures)
  
  #generate random mutation samples of Signature 10
  a <- sample(rownames(sig), size = n, replace = TRUE, prob = as.vector(sig[,"A"]))
  df_a <- as_tibble(a)
  df_a <- df_a %>%
    count(value) %>%
    mutate(
      value = factor(value, levels = order.tri),
      prob = n/sum(n),
    ) %>%
    merge(mut_order, all = TRUE)
  
  df_a[is.na(df_a)] <- 0     # Replace NA with 0
  df_a <- df_a %>%
    mutate(
      prob = prob + 1e-04
    )
  
  #generate random mutation samples of Signature 6
  b <- sample(rownames(sig), size = n, replace = TRUE, prob = as.vector(sig[,"B"]))
  df_b <- as_tibble(b)
  df_b <- df_b %>%
    count(value) %>%
    mutate(
      value = factor(value, levels = order.tri),
      prob = n/sum(n)
    ) %>%
    merge(mut_order, all = TRUE)
  
  df_b[is.na(df_b)] <- 0     # Replace NA with 0
  df_b <- df_b %>%
    mutate(
      prob = prob + 1e-04
    )
  
  #join those two sample mutations
  mut_df <- df_a %>%
    inner_join(df_b, by = "value") %>%
    mutate(
      prob = (prob.x + prob.y)/2
    ) %>%
    select(value, prob)
  
  #define the sample profile created by two sample mutations
  sample_profile <- mut_df %>%
    select(prob)
  sample_profile <- as.matrix(sample_profile)
  
  sample_profile
}
```

```{r}
#A, B = Signatures.A, Signatures.B
#n = the number of mutations that we want to generate
generate_sample_prof(10,6,500)
```

