---
title: "curi1"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library(tidyverse)
library(MutationalPatterns)
library(NMF)
library(reshape2)
library(grid)
library(gridExtra)
library(ggpubr)
library(ggrepel)
library(broom)
library(emplik)
library(glmc)
library(ConsReg)
library(Hmisc)
library(yardstick)

```



## Preliminary functions

The following functions allow us to load and analyze mutational signatures

```{r}
# Loads a mutational signature matrix from a mutational signature txt file
load_old_mutational_matrix<- 
  function(filename,
          mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv") {
  mutational.signature <-  read_csv(filename)
  mut_mat <-  read_csv(mut.order)

  mut.join <-  mutational.signature %>%
    inner_join(mut_mat)
  
  order.tri <-  unique(mut.join$ext.context)

  mut.join <-  mut.join %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    select(ext.context,prob)
  
  mut.matrix <-  as.matrix(
        sapply(mut.join[,2:ncol(mut.join)], as.numeric))  

  rownames(mut.matrix) <-  mut.join$ext.context
 
  return (t(mut.matrix)[,1:96])
}

change_old_mutational_matrix<-
    function(sig.frame,
             mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv") {
        
  mutational.signature <-  sig.frame
  mut_mat <-  read_csv(mut.order)

  mut.join <-  mutational.signature %>%
    inner_join(mut_mat)
  
  order.tri <-  unique(mut.join$ext.context)

  mut.join <-  mut.join %>%
      mutate(ext.context =
                 factor (ext.context, levels = order.tri)) %>%
      select(sample,ext.context,prob) %>%
      pivot_wider(names_from = ext.context,
                  values_from = prob)
    
    mut.matrix <-  as.matrix(
        sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
    rownames(mut.matrix) = mut.join$sample

    return (t(mut.matrix))
}

load_cosmic_matrix<- function(mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv") {

  sp_url <- paste("https://cancer.sanger.ac.uk/cancergenome/assets/",
                "signatures_probabilities.txt", sep = "")
   
  cancer_signatures <-  read.table(sp_url, sep = "\t", header = TRUE)
  # Match the order of the mutation types to MutationalPatterns standard
  mut_mat <-  read_csv(mut.order)
  new_order <-  match(mut_mat$ext.context, cancer_signatures$Somatic.Mutation.Type) 
  # Reorder cancer signatures dataframe
  cancer_signatures <-  cancer_signatures[as.vector(new_order),]
  # Add trinucletiode changes names as row.names
  row.names(cancer_signatures) <-  cancer_signatures$Somatic.Mutation.Type
  # Keep only 96 contributions of the signatures in matrix
  cancer_signatures <- as.matrix(cancer_signatures[,4:33])
  return(cancer_signatures)
}

# Returns a frame with columns Signature, Sample, Contribution with the absolute
# contribution of each signature for each sample
calc_abs_contribution <- function(nmf.res,sig.names) {

  total_signatures <-  colSums(nmf.res$signatures) 
  abs_contribution <-  nmf.res$contribution * total_signatures
  rownames(abs_contribution) <- sig.names

  m_contribution <-  melt(abs_contribution)
  colnames(m_contribution) = c("signature", "sample", "contribution")

  return (m_contribution)
}

# Returns a frames with the cosine similarity (fit) between the original profile and the
# reconstructed profile
calc_fit_contribution <- function(mut.matrix,nmf.res) {

  # calculate all pairwise cosine similarities
  cos_sim_ori_rec <- cos_sim_matrix(mut.matrix, nmf.res$reconstructed)
  # extract cosine similarities per sample between original and reconstructed
  fit.table<- as.data.frame(diag(cos_sim_ori_rec))
  fit.table$Sample <-  rownames(fit.table)
  colnames(fit.table) = c("fit","sample")

  return(fit.table)
}

get_top_signature <-function(mut.table, signatures.table,
                             signature,top.n=8,
                             signatures=
                                 c("PolE","Aging","MMR","Germline")){

    # Construct table with only contribution from specific signature
    sig.table <-  mut.table %>%
        gather(signatures,key="sign.name",value="contribution") %>%
        filter(sign.name==signature) %>%
        arrange(desc(contribution))

    # Pick top n samples for that signature
    top.table <- 
        sig.table %>% arrange(desc(contribution)) %>%
        head(top.n)

    # Pick signatures for those top samples and plot them
    mut.selected <- top.table %>%
        select(sample) %>%
        inner_join (signatures.table)

    top.signatures <-  change_old_mutational_matrix(mut.selected)
    colnames(top.signatures) <-  paste0("P",1:top.n,"")

    plot.top <-  plot_96_profile(top.signatures)

    return (plot.top)
}
```

## Data acquisition and cleaning

We load the mutational signatures file for all of the samples and consider only 
samples that have high sequencing coverage (over 5Mb at 20x)
```{r}
# Load mutational signature data for all samples
signatures <- read_csv("mutational_sig_oec-main/supplied_data/mutational.sig.csv")

# Load sequencing/cohort info for samples, filter with at
# at least 5 million Mb covered
sample.info <-
    read_csv("mutational_sig_oec-main/supplied_data/sample.info.csv") %>%
    filter (area.cov > 5e6)
```

## Linear regression model
```{r}
#Load the P3, Sig10,6,5 data
data <- read_csv("Data/P3_sig.csv")

model1 <- lm(p3 ~ 0 + sig10 + sig6, data = data)
summary(model1)

model2 <- lm(p3 ~ 0 + sig10 + sig1, data = data)
summary(model2)

#plot1
ggplot(data, aes(sig10, p3)) +
  geom_point() +
  geom_smooth(method = "lm")

data1 <- data %>%
  pivot_longer(
    cols = c(`sig10`, `sig6`, `sig1`),
    names_to = "signatures", 
    values_to = "prob"
  )

#plot2
ggplot(data1, aes(x = prob, y = p3, color = signatures)) +
  geom_point() +
  geom_smooth(method = "lm")
```

## Function testing code
```{r}
#load the Signatures 10, 6, 1
cosmic.signatures <- load_cosmic_matrix()
sig <- cosmic.signatures[,c(10,6,1)]
sig.labels <-  c("10","6","1")
colnames(sig) <- sig.labels

#load the mutation order
mut.selected <- sample.info %>%
  select (sample) %>%
  inner_join (signatures)

mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv"
mutational.signature <-  mut.selected
mut_mat <-  read_csv(mut.order)

mut.join <-  mutational.signature %>%
  inner_join(mut_mat)
order.tri <-  unique(mut.join$ext.context)
mut_order <- as_tibble(order.tri)

#generate random mutation samples of Signature 10
a <- sample(rownames(sig), size = 500, replace = TRUE, prob = as.vector(sig[,"10"]))
df_a <- as_tibble(a)
df_a <- df_a %>%
  count(value) %>%
  mutate(
    value = factor(value, levels = order.tri),
    prob = n/sum(n),
  ) %>%
  merge(mut_order, all = TRUE)

df_a[is.na(df_a)] <- 0     # Replace NA with 0
df_a <- df_a %>%
  mutate(
    prob = prob + 1e-04
  )

#generate random mutation samples of Signature 6
b <- sample(rownames(sig), size = 500, replace = TRUE, prob = as.vector(sig[,"6"]))
df_b <- as_tibble(b)
df_b <- df_b %>%
  count(value) %>%
  mutate(
    value = factor(value, levels = order.tri),
    prob = n/sum(n)
  ) %>%
  merge(mut_order, all = TRUE)

df_b[is.na(df_b)] <- 0     # Replace NA with 0
df_b <- df_b %>%
  mutate(
    prob = prob + 1e-04
  )

#join those two sample mutations
mut_df <- df_a %>%
  inner_join(df_b, by = "value") %>%
  mutate(
    prob = (prob.x + prob.y)/2
  ) %>%
  select(value, prob)

#define the sample profile created by two sample mutations
sample_profile <- mut_df %>%
  select(prob)
sample_profile <- as.matrix(sample_profile)
sample_profile
```

## Function for generating the sample mutation profiles
Load the Mutation signatures order
```{r}
mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv"
mutational.signature <-  mut.selected
mut_mat <-  read_csv(mut.order)

mut.join <-  mutational.signature %>%
  inner_join(mut_mat)
order.tri <-  unique(mut.join$ext.context)
mut_order <- as_tibble(order.tri)
```

Function for generating the sample mutation profiles 
```{r}
#A, B = Signatures.A, Signatures.B
#n = the number of mutations that we want to generate
generate_sample_prof <- function(A, B, n){

  cosmic.signatures <- load_cosmic_matrix()
  sig <- cosmic.signatures[,c(A, B)]
  sig.labels <-  c("A","B")
  colnames(sig) <- sig.labels
  
  #load the mutation order
  mut.selected <- sample.info %>%
    select (sample) %>%
    inner_join (signatures)
  
  mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv"
mutational.signature <-  mut.selected
mut_mat <-  read_csv(mut.order)

mut.join <-  mutational.signature %>%
  inner_join(mut_mat)
order.tri <-  unique(mut.join$ext.context)
mut_order <- as_tibble(order.tri)
  
  #generate random mutation samples of Signature 10
  a <- sample(rownames(sig), size = n, replace = TRUE, prob = as.vector(sig[,"A"]))
  df_a <- as_tibble(a)
  df_a <- df_a %>%
    count(value) %>%
    mutate(
      value = factor(value, levels = order.tri),
      prob = n/sum(n),
    ) %>%
    merge(mut_order, all = TRUE)
  
  df_a[is.na(df_a)] <- 0     # Replace NA with 0
  df_a <- df_a %>%
    mutate(
      prob = prob + 1e-04
    )
  
  #generate random mutation samples of Signature 6
  b <- sample(rownames(sig), size = n, replace = TRUE, prob = as.vector(sig[,"B"]))
  df_b <- as_tibble(b)
  df_b <- df_b %>%
    count(value) %>%
    mutate(
      value = factor(value, levels = order.tri),
      prob = n/sum(n)
    ) %>%
    merge(mut_order, all = TRUE)
  
  df_b[is.na(df_b)] <- 0     # Replace NA with 0
  df_b <- df_b %>%
    mutate(
      prob = prob + 1e-04
    )
  
  #join those two sample mutations
  mut_df <- df_a %>%
    inner_join(df_b, by = "value") %>%
    mutate(
      prob = (prob.x + prob.y)/2
    ) %>%
    select(value, prob)
  
  #define the sample profile created by two sample mutations
  sample_profile <- mut_df %>%
    select(prob)
  sample_profile <- as.matrix(sample_profile)
  
  sample_profile
}
```


```{r}
generate_sample_prof_1 <- function(A, B, n1, n2){
  cosmic.signatures <- load_cosmic_matrix()
  sig <- cosmic.signatures[,c(A, B)]
  sig.labels <-  c("A","B")
  colnames(sig) <- sig.labels
  
  #generate random mutation samples of Signature 10
  a <- sample(rownames(sig), size = n1, replace = TRUE, prob = as.vector(sig[,"A"]))
  df_a <- as_tibble(a)
  df_a <- df_a %>%
    count(value) %>%
    mutate(
      value = factor(value, levels = order.tri),
      prob = n/sum(n),
    ) %>%
    merge(mut_order, all = TRUE)
  
  df_a[is.na(df_a)] <- 0     # Replace NA with 0
  
  #generate random mutation samples of Signature 6
  b <- sample(rownames(sig), size = n2, replace = TRUE, prob = as.vector(sig[,"B"]))
  df_b <- as_tibble(b)
  df_b <- df_b %>%
    count(value) %>%
    mutate(
      value = factor(value, levels = order.tri),
      prob = n/sum(n)
    ) %>%
    merge(mut_order, all = TRUE)
  
  df_b[is.na(df_b)] <- 0     # Replace NA with 0
  
  #join those two sample mutations
  mut_df <- df_a %>%
    inner_join(df_b, by = "value") %>%
    mutate(
      prob = ((prob.x * (n1/(n1+n2))) + (prob.y * (n2/(n1+n2))))
    ) %>%
    select(value, prob)
  
  #define the sample profile created by two sample mutations
  sample_profile <- mut_df %>%
    select(prob)
  sample_profile <- as.matrix(sample_profile)
  
  sample_profile
}
```



```{r}
#A, B = Signatures.A, Signatures.B
#n = the number of mutations that we want to generate
generate_sample_prof(10,6,500)
generate_sample_prof(10,1,500)
```

Function for Linear Regression
```{r}
lm_test <- function(sample_profile, sigA, sigB){
  sig <-cosmic.signatures[,c(sigA,sigB)]
  sig.labels <- c("sigA", "sigB")
  colnames(sig) <- sig.labels
 
  data<- cbind(sample_profile,sig)
  data<-as_tibble(data)
  model<- lm(sample_profile ~ 0 + sigA + sigB, data=data)
  
  tidy <- model %>% tidy()
  tidy$estimate
}

lm_test(data$p3,10,6)
```

Testing LR Function with Simulated Data
```{r}
test_data <- generate_sample_prof(10,6,500)

function_test(test_data, 10, 6)
#output is expected with each coefficient being 0.5 or 1/2, since this is what we put into the sample mutational profile called test_data
```

Modifying the function to create profiles with different number of mutations
```{r}
generate_sample_prof <- function(A, B, n1, n2){

  cosmic.signatures <- load_cosmic_matrix()
  sig <- cosmic.signatures[,c(A, B)]
  sig.labels <-  c("A","B")
  colnames(sig) <- sig.labels
  
  #generate random mutation samples of Signature 10
  a <- sample(rownames(sig), size = n1, replace = TRUE, prob = as.vector(sig[,"A"]))
  df_a <- as_tibble(a)
  df_a <- df_a %>%
    count(value) %>%
    mutate(
      value = factor(value, levels = order.tri),
      prob = n/sum(n),
    ) %>%
    merge(mut_order, all = TRUE)
  
  df_a[is.na(df_a)] <- 0     # Replace NA with 0
  
  #generate random mutation samples of Signature 6
  b <- sample(rownames(sig), size = n2, replace = TRUE, prob = as.vector(sig[,"B"]))
  df_b <- as_tibble(b)
  df_b <- df_b %>%
    count(value) %>%
    mutate(
      value = factor(value, levels = order.tri),
      prob = n/sum(n)
    ) %>%
    merge(mut_order, all = TRUE)
  
  df_b[is.na(df_b)] <- 0     # Replace NA with 0
  
  #join those two sample mutations
  mut_df <- df_a %>%
    inner_join(df_b, by = "value") %>%
    mutate(
      prob = ((prob.x * (n1/(n1+n2))) + (prob.y * (n2/(n1+n2))))
    ) %>%
    select(value, prob)
  
  #define the sample profile created by two sample mutations
  sample_profile <- mut_df %>%
    select(prob)
  sample_profile <- as.matrix(sample_profile)
  
  sample_profile
}
```

Testing Out Updated Function (disregard)
```{r}
test_data2 <- generate_sample_prof_new(10, 6, 600, 400)
function_test(test_data2, 10, 6)
#this is not as accurate, we would expect the coefficients to be 0.6 for Sig 10 (y) and 0.4 for Sig 6 (z)

test_data3 <- generate_sample_prof_new(10, 6, 900, 100)
function_test(test_data3, 10, 6)
#again, not ideal results. going back to make sure I modified the function correctly
#note: everything looks correct in the function

test_data4 <- generate_sample_prof_new(1, 4, 900, 100)
function_test(test_data4, 1, 4)
#coefficient are quite different near what we would expect

test_data4 <- generate_sample_prof_new(1, 4, 500, 500)
function_test(test_data4, 1, 4)
#coefficients are close, but still not accurate to our simulated data

test_data5 <- generate_sample_prof_new(1, 10, 1000, 9000)
function_test(test_data5, 1, 10)
#coefficient are quite different near what we would expect

test_data6 <- generate_sample_prof_new(1, 10, 1, 9)
function_test(test_data6, 1, 10)
#coefficient are quite different near what we would expect

test_data7 <- generate_sample_prof_new(1, 10, 3000, 7000)
function_test(test_data7, 1, 10)
#coefficient are quite different near what we would expect
```


fit_to_signatures exploration
```{r}
sample1 <- generate_sample_prof(10, 6, 600, 400)

signatures <- get_known_signatures()

#Test with simulated sample profile
fit_to_signatures(sample1, signatures)
fit_test <- fit_to_signatures_strict(sample1, signatures[,c(6, 13, 14)], max_delta = 0.002)
fit_test <- fit_test$fit_res$contribution
df <- cbind(signatures = rownames(fit_test), fit_test)
rownames(df) <- 1:nrow(df)
df <- as_tibble(df) %>%
  pivot_wider(names_from = signatures, values_from = prob)
df %>%
  mutate(SBS10 = as.numeric(SBS10a) + as.numeric(SBS10b), 
         SBS6 = as.numeric(SBS6))

#Another simulated profile
sample2 <- generate_sample_prof(6, 11, 100, 900)
fit_to_signatures(sample2, signatures)
fit_test <- fit_to_signatures_strict(sample2, signatures[,c(6, 17)], max_delta = 0.002)
fit_test <- fit_test$fit_res$contribution
df <- cbind(signatures = rownames(fit_test), fit_test)
rownames(df) <- 1:nrow(df)
df <- as_tibble(df) %>%
  pivot_wider(names_from = signatures, values_from = prob)
df %>%
  mutate(SBS11 = as.numeric(SBS11), 
         SBS6 = as.numeric(SBS6))

#Testing function with P3 sample
P3 <- data$p3
P3 <- as.matrix(P3)
fit_to_signatures(P3, signatures)
fit_test1 <- fit_to_signatures_strict(P3, signatures[,c(1, 6, 13, 14)], max_delta = 0.002)
fit_test1 <- fit_test1$fit_res$contribution
df <- cbind(signatures = rownames(fit_test1), fit_test1)
rownames(df) <- 1:nrow(df)
df <- as_tibble(df) %>%
  pivot_wider(names_from = signatures, values_from = "1")
df %>%
  mutate(SBS10 = as.numeric(SBS10a) + as.numeric(SBS10b), 
         SBS6 = as.numeric(SBS6))

```

Testing fit_to_signatures against our function
```{r}
#Comparision 1
sample2 <- generate_sample_prof(6, 11, 100, 900)
fit_to_signatures(sample2, signatures)
fit_test <- fit_to_signatures_strict(sample2, signatures[,c(6, 17)], max_delta = 0.002)
fit_test <- fit_test$fit_res$contribution
df <- cbind(signatures = rownames(fit_test), fit_test)
rownames(df) <- 1:nrow(df)
df <- as_tibble(df) %>%
  pivot_wider(names_from = signatures, values_from = prob)
df %>%
  mutate(SBS11 = as.numeric(SBS11), 
         SBS6 = as.numeric(SBS6))

lm_test(sample2, 6, 11)

#Comparision 2
sample3 <- generate_sample_prof(1, 4, 500, 500)
fit_to_signatures(sample3, signatures)
fit_test <- fit_to_signatures_strict(sample3, signatures[,c(1, 4)], max_delta = 0.002)
fit_test <- fit_test$fit_res$contribution
df <- cbind(signatures = rownames(fit_test), fit_test)
rownames(df) <- 1:nrow(df)
df <- as_tibble(df) %>%
  pivot_wider(names_from = signatures, values_from = prob)
df %>%
  mutate(SBS1 = as.numeric(SBS1), 
         SBS4 = as.numeric(SBS4))

lm_test(sample3, 1, 4)

#Comparision 3
sample4 <- generate_sample_prof(1, 6, 200, 800)
fit_to_signatures(sample4, signatures)
fit_test <- fit_to_signatures_strict(sample4, signatures[,c(1, 6)], max_delta = 0.001)
fit_test <- fit_test$fit_res$contribution
df <- cbind(signatures = rownames(fit_test), fit_test)
rownames(df) <- 1:nrow(df)
df <- as_tibble(df) %>%
  pivot_wider(names_from = signatures, values_from = prob)
df %>%
  mutate(SBS1 = as.numeric(SBS1), 
         SBS6 = as.numeric(SBS6))

lm_test(sample4, 1, 6)
```


```{r}


fit2 = ConsReg(formula = p3 ~ 0 + sig10 + sig6, data = data,
            family = 'gaussian',
            constraints = '0.08 <sig10< 0.152, 0.2<sig6 < 0.02',
            LOWER = 0.001, UPPER = 0.1,
            )
fit2


```

Function for Linear Regression
```{r}
cosmic.signatures <- get_known_signatures()
sig <-cosmic.signatures[,c("SBS6","SBS10a")]
sig.labels <- c("sigA", "sigB")
colnames(sig) <- sig.labels

data<- cbind(sample_profile,sig)
data<-as_tibble(data)

constraints <- str_c(0.1, ' < sigA, sigA < ', 1,', ', 0.1, ' < sigB, sigB < ', 1)

model<- ConsReg(formula = prob ~ 0 + sigA + sigB, data = data,
          family = 'gaussian',
          constraints = constraints
          )

coeff <- as.tibble(model$coefficients)
coeff$value
```

## testing
```{r}
sample_profile <- generate_sample_prof(1, 6, 200, 800)
```

```{r}
lm_test_2 <- function(sample_profile, sigA, sigB, a1, a2, b1, b2){
  cosmic.signatures <- get_known_signatures()
  sig <-cosmic.signatures[,c(sigA,sigB)]
  sig.labels <- c("sigA", "sigB")
  colnames(sig) <- sig.labels
  
  df <- cbind(sample_profile,sig)
  df <- as_tibble(df)
  
  constraints <- str_c(a1, ' < sigA, sigA < ', a2,', ', b1, ' < sigB, sigB < ', b2)
  
  model<- ConsReg(prob ~ 0 + sigA + sigB, data = df,
            family = 'gaussian',
            constraints = constraints
            )
  
  model$coefficients
  #coeff <- as.tibble(model$coefficients)
  #coeff$value
}

lm_test_2(sample_profile,"SBS10a","SBS6", 0.1, 1, 0.1, 1)
```

## function
```{r}
lm_test_2 <- function(sample_profile, sigA, sigB, a1, a2, b1, b2){
  cosmic.signatures <- get_known_signatures()
  sig <-cosmic.signatures[,c(sigA,sigB)]
  sig.labels <- c("sigA", "sigB")
  colnames(sig) <- sig.labels
  
  df <- cbind(sample_profile,sig)
  df <- as_tibble(df)
  constraint <- str_c(a1, ' < sigA, sigA < ', a2, b1, ' < sigB, sigB < ', b2)
  
  model<- ConsReg(prob ~ 0 + sigA + sigB, data = df,
            family = 'gaussian',
            constraints = 'a1 < sigA, sigA < a2, b1 < sigB, sigB < b2'
            )
  
  model$coefficients
  #coeff <- as.tibble(model$coefficients)
  #coeff$value
}

lm_test_2(sample_profile,"SBS10a","SBS6", 0.1, 1, 0.1, 1)
```


```{r}
lm_test <- function(sample_profile, sigs, constrs){
  
  cosmic.signatures <- get_known_signatures()
  sig <- cosmic.signatures[,sigs]
  
  df <- cbind(sample_profile,sig)
  df <- as_tibble(df)
  total_mut <- 1500
  total_size <- 50
  
  constraints <- str_c(constrs[1], ' < ', ((sigs[1]*total_mut)/total_size), ', ', ((sigs[1]*total_mut)/total_size), ' < ', constrs[2])
  for (i in 2:length(sigs)){
    constraint <- str_c(constrs[2*i-1], ' < ', ((sigs[i]*total_mut)/total_size), ', ', ((sigs[i]*total_mut)/total_size), ' < ', constrs[2*i])
    constraints <- str_c(constraints, ', ', constraint)
  }
  
  model<- ConsReg(formula = prob ~ 0 + ., data = df,
            family = 'gaussian',
            constraints = constraints
            )
  
  model$coefficients
  coeff <- as.tibble(model$coefficients)
  coeff$value
}
```


```{r}
sample_profile <- generate_sample_prof(10,6,4000,6000)
lm_test(sample_profile,10,6)
```


```{r}
constraints_ds <- read.csv("~/Desktop/curi2022/Data/constraints_ds.csv")
```




```{r}
estimate_coeff <- function(sample_profile, sigs, tissue_type, scale, quantiles, mb_size){
  
  #parameters testing
  #sample_profile <- generate_sample_prof(c("SBS1","SBS10b","SBS6"),c(100,300,600))
  total_mut <- unname(colSums(sample_profile)[1])
  
  sample_profile <- as.matrix(sample_profile[,2])
  colnames(sample_profile) <- c("prob")
  
  #sigs <- c("SBS1","SBS10b","SBS6")
  #tissue_type <- "BRCA"
  #scale <- "exome"
  #quantiles <- c("5%","95%")
  #mb_size <- 50
  
  #pulling constraints
  constraints_ds <- read_csv("~/Desktop/curi2022/Data/constraints_ds.csv")
  constr1 <- constraints_ds %>%
    filter(Tissue == tissue_type, Signature %in% sigs, Scale == scale, Quantile %in% quantiles) 
  constrs <- c(constr1$Value)
  
  #making dataset for the model
  cosmic.signatures <- get_known_signatures()
  sig <- cosmic.signatures[,sigs]
  
  df <- cbind(sample_profile,sig)
  df <- as_tibble(df)
  
  constraints <- str_c(constrs[1]*(mb_size/total_mut), ' < ', sigs[1], ', ', sigs[1], ' < ', constrs[2]*(mb_size/total_mut))
  for (i in 2:length(sigs)){
    constraint <- str_c(constrs[2*i-1]*(mb_size/total_mut), ' < ', sigs[i], ', ', sigs[i], ' < ', constrs[2*i]*(mb_size/total_mut))
    constraints <- str_c(constraints, ', ', constraint)
  }
  
  model<- ConsReg(formula = prob ~ 0 + ., data = df,
            family = 'gaussian',
            constraints = constraints
            )
  
  model$coefficients
  #coeff <- as.tibble(model$coefficients)
  #coeff$value
}
```


```{r}
sample_profile <- generate_sample_prof(c("SBS10a","SBS10b","SBS1","SBS6", "SBS2"),c(200,300,100,300,100))
plot_sample_prof(sample_profile)
estimate_coeff(sample_profile, sigs = c("SBS10a","SBS10b","SBS1","SBS6", "SBS2"), tissue_type = "BRCA", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


sample_profile <- generate_sample_prof(c("SBS10a","SBS10b","SBS1","SBS6"),c(300,100,200,400))
plot_sample_prof(sample_profile)
estimate_coeff(sample_profile, sigs = c("SBS10a","SBS10b","SBS1","SBS6"), tissue_type = "BRCA", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)
```

For colorectal cancer

```{r}

```


```{r}
constraints_ds %>%
  count(Tissue)


sample_profile <- generate_sample_prof(c("SBS1","SBS6"),c(400,600))
plot_sample_prof(sample_profile)
estimate_coeff(sample_profile, sigs = c("SBS1","SBS6"), tissue_type = "CHOL", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)  
```

```{r}
#loading in COSMIC sample data
COSMICsample_ds <- read_csv("Data/COSMICsample_ds.csv")

#getting just colorectal cancer samples
colo_samples <- COSMICsample_ds %>%
  filter(Cancer.Types == "ColoRect-AdenoCA")
read(colo_samples)

#creating sample SP18310
sample_prof <- generate_sample_prof(c("SBS1","SBS5","SBS44"),c(33221,25724,49282))
test1 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5","SBS44"), tissue_type = "COAD", scale = "genome", quantiles = c("5%", "95%"), mb_size = 3000)


#creating sample SP17172
sample_prof <- generate_sample_prof(c("SBS1","SBS5","SBS44"),c(31717,46279,46366))
test2 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5","SBS44"), tissue_type = "COAD", scale = "genome", quantiles = c("5%", "95%"), mb_size = 3000)

actual <- c(test1[1], test1[2], test1[3], test2[1], test2[2], test2[3])
expected <- c(33221, 25724, 49282, 31717,46279,46366)
data.frame(actual, expected)

```

```{r}

#828, sample number TCGA-AA-3549-01A-02D-1953-10
sample_prof <- generate_sample_prof(c("SBS15"),c(13))
proptable828 <- prop.table(c(13))
test828 <- estimate_coeff(sample_prof, sigs = c("SBS15"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)
#error


#827, sample number HT115
sample_prof <- generate_sample_prof(c("SBS5","SBS10a","SBS10b", "SBS15","SBS28"),c(1064,1077,2226,1231,474))
proptable827 <- prop.table(c(1064,1077,2226,1231,474))
test827 <- estimate_coeff(sample_prof, sigs = c("SBS5","SBS10a","SBS10b", "SBS15","SBS28"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


#826, sample number VACO10
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(63,129))
proptable826 <- prop.table(c(63,129))
test826 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#825, sample number T96
sample_prof <- generate_sample_prof(c("SBS1","SBS5","SBS10b"),c(186,86,94))
proptable825 <- prop.table(c(186,86,94))
test825 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5","SBS10b"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#824, sample number T70
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(39,44))
proptable824 <- prop.table(c(39,44))
test824 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#823, sample number T684
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(49,60))
proptable823 <- prop.table(c(49,60))
test823 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


#822, sample number T660
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(39,63))
proptable822 <- prop.table(c(39,63))
test822 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#821, sample number T636
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(123,135))
proptable821 <- prop.table(c(123,135))
test821 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#820, sample number T613
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(215,207))
proptable820 <- prop.table(c(215,207))
test820 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


#819, sample number T595
sample_prof <- generate_sample_prof(c("SBS1"),c(34))
proptable819 <- prop.table(c(34))
test819 <- estimate_coeff(sample_prof, sigs = c("SBS1"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)
#error


#818, sample number T578
sample_prof <- generate_sample_prof(c("SBS1","SBS5", "SBS10a","SBS10b","SBS28"),c(343,383,1663,	1447,546))
proptable818 <- prop.table(c(343,383,1663,1447,546))
test818 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5", "SBS10a","SBS10b","SBS28"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#817, sample number T55
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(76,384))
proptable817 <- prop.table(c(76,384))
test817 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5", "SBS6"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


#816, sample number T521
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(38,45))
proptable816 <- prop.table(c(38,45))
test816 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


#815, sample number T475
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(39,43))
proptable815 <- prop.table(c(39,43))
test815 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


#814, sample number T473
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(27,32))
proptable814 <- prop.table(c(27,32))
test814 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


#813, sample number T469
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(130,85))
proptable813 <- prop.table(c(130,85))
test813 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#812, sample number T464
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(29,17))
proptable812 <- prop.table(c(29,17))
test812 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#811, sample number T407
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(	204,157))
proptable811 <- prop.table(c(204,157))
test811 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#810, sample number T388
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(20,24))
proptable810 <- prop.table(c(20,24))
test810 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#809, sample number T3724
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(432,830))
proptable809 <- prop.table(c(432,830))
test809 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


#808, sample number T3720
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(49,61))
proptable808 <- prop.table(c(49,61))
test808 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


#807, sample number T368
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(306,322))
proptable807 <- prop.table(c(306,322))
test807 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


#806, sample number T3670
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(72,91))
proptable806 <- prop.table(c(72,91))
test806 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


#805, sample number T3668
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(77,91))
proptable805 <- prop.table(c(77,91))
test805 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#804, sample number T3658
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(295,433))
proptable804 <- prop.table(c(295,433))
test804 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


#803, sample number T3610
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(196,71))
proptable803 <- prop.table(c(196,71))
test803 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


#802, sample number T3603
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(62,43))
proptable802 <- prop.table(c(62,43))
test802 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


#801, sample number T36
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(84,128))
proptable801 <- prop.table(c(84,128))
test801 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#800, sample number T3557
sample_prof <- generate_sample_prof(c("SBS1","SBS5"),c(51,54))
proptable800 <- prop.table(c(51,54))
test800 <- estimate_coeff(sample_prof, sigs = c("SBS1","SBS5"), tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)


actual <- c(test826[1], test826[2], test825[1],test825[2],test825[3], test824[1],test824[2],test823[1],test823[2],test822[1],test822[2],test821[1],test821[2],test820[1],test820[2],test818[1],test818[2],test818[3],test818[4],test818[5],test817[1],test817[2],test816[1],test816[2],test815[1],test815[2],test814[1],test814[2],test813[1],test813[2],test812[1],test812[2],test811[1],test811[2],test810[1],test810[2],test809[1],test809[2],test808[1],test808[2],test807[1],test807[2],test806[1],test806[2],test805[1],test805[2],test804[1],test804[2],test803[1],test803[2],test802[1],test802[2],test801[1],test801[2],test800[1],test800[2])


expected<-c( proptable826,proptable825,proptable824,proptable823,proptable822,proptable821,proptable820,proptable818,proptable817,proptable816,proptable815,proptable814,proptable813,proptable812,proptable811,proptable810,proptable809,proptable808,proptable807,proptable806,proptable805,proptable804,proptable803,proptable802,proptable801,proptable800 )

graph<-data.frame(actual,expected)



```




Loop for comparing actual and method results

```{r}
COSMICsample_ds <- read_csv("Data/COSMICsample_ds.csv")
cosmic.signatures <- get_known_signatures()
tissue_list <- unique(COSMICsample_ds$Cancer.Types)
```

```{r}
COSMICsample <- COSMICsample_ds %>%
  select(6:64) 
signature_list <- colnames(COSMICsample)
signature_list <- str_remove(signature_list, ".Mutations")
cosmic_signature_list <- colnames(cosmic.signatures)

known_sig <- signature_list[signature_list %in% cosmic_signature_list]
```

```{r}
coloE_samples <- COSMICsample_ds %>%
  filter(Cancer.Types == "ColoRect-AdenoCA") %>%
  filter(Scale == "exome")

coloE_samples <- coloE_samples %>%
  select(4,6:64) 
coloE_size <- dim(coloE_samples)[1]

actual_list <- c()
expected_list <- c()
sample_names <- c()
for (i in 1:coloE_size){
  mut <- as.numeric(coloE_samples[i,])
  nonzero_sig <- (which(mut %nin% c(0, NA)))-1
  coloE_samples_noname <- coloE_samples %>%
    select(2:60)
  list <- colnames(coloE_samples_noname)
  s_list <- list[nonzero_sig]
  s_list <- str_remove(s_list, ".Mutations")
  m_list <- mut[mut %nin% c(0, NA)]
  
  s_list <- s_list[s_list %in% known_sig]
  m_list <- m_list[s_list %in% known_sig]
  
  sample_prof <- generate_sample_prof(s_list, m_list)
  prop.test <- prop.table(m_list)
  method_test <- estimate_coeff(sample_prof, sigs = s_list, tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)
  
  for(j in 1:length(s_list)){
    expected <- prop.test
    actual <- method_test
    sample_names_list <- coloE_samples %>%
      select(1)
    sample_names_list <- as.matrix(sample_names_list)
    sample_name <- sample_names_list[i]
    sample_names <- append(sample_names, sample_name)
  }
  actual_list <- append(actual_list, actual)
expected_list <- append(expected_list, expected)
df_1 <- cbind(actual_list, expected_list)
  df_1 <- as_tibble(df_1)
  mse <- rmse_vec(df_1$actual_list, df_1$expected_list)^2
  mse_list <- append(mse_list, mse)

}

df <- cbind(expected_list, actual_list, sample_names)
  df <- as_tibble(df)
  

```


Testing with one sample with MSE
```{r}
coloE_samples <- COSMICsample_ds %>%
  filter(Cancer.Types == "ColoRect-AdenoCA") %>%
  filter(Scale == "exome")

coloE_samples <- coloE_samples %>%
  select(4,6:64) 
coloE_size <- dim(coloE_samples)[1]
actual_list <- c()
expected_list <- c()
sample_names <- c()
mse_list <- c()

mut <- as.numeric(coloE_samples[1,])
  nonzero_sig <- (which(mut %nin% c(0, NA)))-1
  coloE_samples_noname <- coloE_samples %>%
    select(2:60)
  list <- colnames(coloE_samples_noname)
  s_list <- list[nonzero_sig]
  s_list <- str_remove(s_list, ".Mutations")
  m_list <- mut[mut %nin% c(0, NA)]
  
  s_list <- s_list[s_list %in% known_sig]
  m_list <- m_list[s_list %in% known_sig]
  
  sample_prof <- generate_sample_prof(s_list, m_list)
  prop.test <- prop.table(m_list)
  method_test <- estimate_coeff(sample_prof, sigs = s_list, tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)
  
  for(j in 1:length(s_list)){
    expected <- prop.test
    actual <- method_test
    sample_names_list <- coloE_samples %>%
      select(1)
    sample_names_list <- as.matrix(sample_names_list)
    sample_name <- sample_names_list[1]
    sample_names <- append(sample_names, sample_name)
  
actual_list <- append(actual_list, actual)
expected_list <- append(expected_list, expected)
df_1 <- cbind(actual_list, expected_list, sample_names)
  df_1 <- as_tibble(df_1)
  mse <- rmse_vec(df_1$actual_list, df_1$expected_list)^2
  mse_list <- append(mse_list, mse)
  }
mse_list
```



```{r}
#Testing for all samples with mse

actual_list <- c()
expected_list <- c()
sample_names <- c()

for (i in 1:coloE_size){
  mut <- as.numeric(coloE_samples[i,])
  nonzero_sig <- (which(mut %nin% c(0, NA)))-1
  coloE_samples_noname <- coloE_samples %>%
    select(2:60)
  list <- colnames(coloE_samples_noname)
  s_list <- list[nonzero_sig]
  s_list <- str_remove(s_list, ".Mutations")
  m_list <- mut[mut %nin% c(0, NA)]
  
  s_list <- s_list[s_list %in% known_sig]
  m_list <- m_list[s_list %in% known_sig]
  
  sample_prof <- generate_sample_prof(s_list, m_list)
  prop.test <- prop.table(m_list)
  method_test <- estimate_coeff(sample_prof, sigs = s_list, tissue_type = "COAD", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)
  
  for(j in 1:length(s_list)){
    expected <- prop.test
    actual <- method_test
    sample_names_list <- coloE_samples %>%
      select(1)
    sample_names_list <- as.matrix(sample_names_list)
    sample_name <- sample_names_list[i]
    sample_names <- append(sample_names, sample_name)
  
actual_list <- append(actual_list, actual)
expected_list <- append(expected_list, expected)
df_1 <- cbind(actual_list, expected_list, sample_names)
  df_1 <- as_tibble(df_1)
  mse <- rmse_vec(df_1$actual_list, df_1$expected_list)^2
  mse_list <- append(mse_list, mse)
  }}
mse_list

```


Graph without mse

```{r}
#reading in comparison-ds

comparison_ds <- read_csv("~/Desktop/curi2022/Data/comparison_ds.csv")
View(comparison_ds)



ggplot(data = comparision_data, mapping = aes(x = expected_list, y = actual_list))+
   geom_point(alpha=0.3)+
   geom_smooth()
 
 
 
 
comparision_data %>%
  filter(actual_list > 1)
 
df_1 %>%
  filter(sample_names == "TCGA-AA-3524-01A-02D-1953-10")
 
                                
 ggplot(data = comparision_data, mapping = aes(x = expected_list, y = actual_list))+
   geom_point(aes(color=actual_list>0.2))+
   geom_smooth()

  

ggplot(comparision_data, aes(x = expected_list, y = actual_list, label= sample_names)) +
  geom_point(data = comparision_data[comparision_data$actual_list>1,], color = "grey50") +
  geom_text_repel(box.padding = 0.5, max.overlaps = Inf) +
  geom_point(data = comparision_data[comparision_data$actual_list<1,], color = "red")



```
