---
title: "Curi Draft"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library(tidyverse)
library(MutationalPatterns)
library(NMF)
library(reshape2)
library(grid)
library(gridExtra)
library(ggpubr)
library(ggrepel)
library(broom)
library(emplik)
library(glmc)
library(ConsReg)
```

## Preliminary functions

The following functions allow us to load and analyze mutational signatures

```{r}
# Loads a mutational signature matrix from a mutational signature txt file
load_old_mutational_matrix<- 
  function(filename,
          mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv") {
  mutational.signature <-  read_csv(filename)
  mut_mat <-  read_csv(mut.order)

  mut.join <-  mutational.signature %>%
    inner_join(mut_mat)
  
  order.tri <-  unique(mut.join$ext.context)

  mut.join <-  mut.join %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    select(ext.context,prob)
  
  mut.matrix <-  as.matrix(
        sapply(mut.join[,2:ncol(mut.join)], as.numeric))  

  rownames(mut.matrix) <-  mut.join$ext.context
 
  return (t(mut.matrix)[,1:96])
}

change_old_mutational_matrix<-
    function(sig.frame,
             mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv") {
        
  mutational.signature <-  sig.frame
  mut_mat <-  read_csv(mut.order)

  mut.join <-  mutational.signature %>%
    inner_join(mut_mat)
  
  order.tri <-  unique(mut.join$ext.context)

  mut.join <-  mut.join %>%
      mutate(ext.context =
                 factor (ext.context, levels = order.tri)) %>%
      select(sample,ext.context,prob) %>%
      pivot_wider(names_from = ext.context,
                  values_from = prob)
    
    mut.matrix <-  as.matrix(
        sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
    rownames(mut.matrix) = mut.join$sample

    return (t(mut.matrix))
}

load_cosmic_matrix<- function(mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv") {

  sp_url <- paste("https://cancer.sanger.ac.uk/cancergenome/assets/",
                "signatures_probabilities.txt", sep = "")
   
  cancer_signatures <-  read.table(sp_url, sep = "\t", header = TRUE)
  # Match the order of the mutation types to MutationalPatterns standard
  mut_mat <-  read_csv(mut.order)
  new_order <-  match(mut_mat$ext.context, cancer_signatures$Somatic.Mutation.Type) 
  # Reorder cancer signatures dataframe
  cancer_signatures <-  cancer_signatures[as.vector(new_order),]
  # Add trinucletiode changes names as row.names
  row.names(cancer_signatures) <-  cancer_signatures$Somatic.Mutation.Type
  # Keep only 96 contributions of the signatures in matrix
  cancer_signatures <- as.matrix(cancer_signatures[,4:33])
  return(cancer_signatures)
}

# Returns a frame with columns Signature, Sample, Contribution with the absolute
# contribution of each signature for each sample
calc_abs_contribution <- function(nmf.res,sig.names) {

  total_signatures <-  colSums(nmf.res$signatures) 
  abs_contribution <-  nmf.res$contribution * total_signatures
  rownames(abs_contribution) <- sig.names

  m_contribution <-  melt(abs_contribution)
  colnames(m_contribution) = c("signature", "sample", "contribution")

  return (m_contribution)
}

# Returns a frames with the cosine similarity (fit) between the original profile and the
# reconstructed profile
calc_fit_contribution <- function(mut.matrix,nmf.res) {

  # calculate all pairwise cosine similarities
  cos_sim_ori_rec <- cos_sim_matrix(mut.matrix, nmf.res$reconstructed)
  # extract cosine similarities per sample between original and reconstructed
  fit.table<- as.data.frame(diag(cos_sim_ori_rec))
  fit.table$Sample <-  rownames(fit.table)
  colnames(fit.table) = c("fit","sample")

  return(fit.table)
}

get_top_signature <-function(mut.table, signatures.table,
                             signature,top.n=8,
                             signatures=
                                 c("PolE","Aging","MMR","Germline")){

    # Construct table with only contribution from specific signature
    sig.table <-  mut.table %>%
        gather(signatures,key="sign.name",value="contribution") %>%
        filter(sign.name==signature) %>%
        arrange(desc(contribution))

    # Pick top n samples for that signature
    top.table <- 
        sig.table %>% arrange(desc(contribution)) %>%
        head(top.n)

    # Pick signatures for those top samples and plot them
    mut.selected <- top.table %>%
        select(sample) %>%
        inner_join (signatures.table)

    top.signatures <-  change_old_mutational_matrix(mut.selected)
    colnames(top.signatures) <-  paste0("P",1:top.n,"")

    plot.top <-  plot_96_profile(top.signatures)

    return (plot.top)
}
```

## Data acquisition and cleaning

We load the mutational signatures file for all of the samples and consider only 
samples that have high sequencing coverage (over 5Mb at 20x)
```{r}
# Load mutational signature data for all samples
signatures <- read_csv(
    "mutational_sig_oec-main/supplied_data/mutational.sig.csv")

# Load sequencing/cohort info for samples, filter with at
# at least 5 million Mb covered
sample.info <-
    read_csv("mutational_sig_oec-main/supplied_data/sample.info.csv") %>%
    filter (area.cov > 5e6)
```

## Linear regression model
This code generates the simple Linear regression model for profile p3 based on Signature 10 and Signature6 or Signature1.
```{r}
#Load the P3, Sig10,6,5 data
data <- read_csv("Data/P3_sig.csv")

model1 <- lm(p3 ~ 0 + sig10 + sig6, data = data)
summary(model1)

model2 <- lm(p3 ~ 0 + sig10 + sig1, data = data)
summary(model2)

#plot1
ggplot(data, aes(sig10, p3)) +
  geom_point() +
  geom_smooth(method = "lm")

data1 <- data %>%
  pivot_longer(
    cols = c(`sig10`, `sig6`, `sig1`),
    names_to = "signatures", 
    values_to = "prob"
  )

#plot2
ggplot(data1, aes(x = prob, y = p3, color = signatures)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    y = "Mutation prob of P3",
    x = "Mutation prob of Sig",
    title = "Contribution of Signatures to the Profile P3"
  ) +
  theme(
        axis.text.x = element_text(color = "grey20", size = 17, angle = 0, hjust = 1, vjust = 0, face = "plain"), axis.text.y = element_text(color = "grey20", size = 17, angle = 0, hjust = 1, vjust = 0, face = "plain"), axis.title = element_text(size = 17), plot.title = element_text(size=22), legend.text=element_text(size=17), legend.title =element_text(size=17)) 
```

## Function testing code
This chuck is for testing the code before making the function.
It is generating the sample mutational profile based on the selected signatures and the number of mutations of each signature.
```{r}
#load the Signatures 10, 6, 1
cosmic.signatures <- load_cosmic_matrix()
sig <- cosmic.signatures[,c(10,6)]

sig.labels <-  c("10","6")
colnames(sig) <- sig.labels

#load the mutation order
mut.selected <- sample.info %>%
  select (sample) %>%
  inner_join (signatures)

mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv"
mutational.signature <-  mut.selected
mut_mat <-  read_csv(mut.order)

mut.join <-  mutational.signature %>%
  inner_join(mut_mat)
order.tri <-  unique(mut.join$ext.context)
mut_order <- as_tibble(order.tri)


#define the number of generating mutations of each signature
n1 <- 5000
n2 <- 5000

#generate random mutation samples of Signature 10
a <- sample(rownames(sig), size = n1, replace = TRUE, prob = as.vector(sig[,"10"]))
df_a <- as_tibble(a)
df_a <- df_a %>%
  count(value) %>%
  mutate(
    value = factor(value, levels = order.tri),
    prob = n/sum(n),
  ) %>%
  merge(mut_order, all = TRUE)

df_a[is.na(df_a)] <- 0     # Replace NA with 0

#generate random mutation samples of Signature 6
b <- sample(rownames(sig), size = n2, replace = TRUE, prob = as.vector(sig[,"6"]))
df_b <- as_tibble(b)
df_b <- df_b %>%
  count(value) %>%
  mutate(
    value = factor(value, levels = order.tri),
    prob = n/sum(n)
  ) %>%
  merge(mut_order, all = TRUE)

df_b[is.na(df_b)] <- 0     # Replace NA with 0

#join those two sample mutations
mut_df <- df_a %>%
  inner_join(df_b, by = "value") %>%
  mutate(
    sum_n = n.x + n.y,
    prob = ((prob.x * (n1/(n1+n2))) + (prob.y * (n2/(n1+n2))))
  ) %>%
  select(value, prob)


#define the sample profile created by two sample mutations
sample_profile <- mut_df %>%
  select(prob)

sample_profile <- as.matrix(sample_profile)
sample_profile


#testing the lm model
sig <-cosmic.signatures[,c(10,6)]
sig.labels <- c("sigA", "sigB")
colnames(sig) <- sig.labels

data<- cbind(sample_profile,sig)
data<-as_tibble(data)
model<- lm(prob ~ 0 + sigA + sigB, data=data)
summary(model)

tidy <- model %>% tidy()
(coefs <- tidy$estimate)
```

## New Cosmic signatures Function testing code
```{r}
#load the Signatures 10, 6, 1
cosmic.signatures <- get_known_signatures()
sig <- cosmic.signatures[,c("SBS7a","SBS7c")]

#load the mutation order
mut.selected <- sample.info %>%
  select (sample) %>%
  inner_join (signatures)

mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv"
mutational.signature <-  mut.selected
mut_mat <-  read_csv(mut.order)

mut.join <-  mutational.signature %>%
  inner_join(mut_mat)
order.tri <-  unique(mut.join$ext.context)

rownames(sig) <- order.tri
mut_order <- as_tibble(order.tri)


#define the number of generating mutations of each signature
n1 <- 5000
n2 <- 5000

#generate random mutation samples of Signature 10
a <- sample(rownames(sig), size = n1, replace = TRUE, prob = as.vector(sig[,"SBS7a"]))
df_a <- as_tibble(a)
df_a <- df_a %>%
  count(value) %>%
  mutate(
    value = factor(value, levels = order.tri),
    prob = n/sum(n),
  ) %>%
  merge(mut_order, all = TRUE)

df_a[is.na(df_a)] <- 0     # Replace NA with 0

#generate random mutation samples of Signature 6
b <- sample(rownames(sig), size = n2, replace = TRUE, prob = as.vector(sig[,"SBS7c"]))
df_b <- as_tibble(b)
df_b <- df_b %>%
  count(value) %>%
  mutate(
    value = factor(value, levels = order.tri),
    prob = n/sum(n)
  ) %>%
  merge(mut_order, all = TRUE)

df_b[is.na(df_b)] <- 0     # Replace NA with 0

#join those two sample mutations
mut_df <- df_a %>%
  inner_join(df_b, by = "value") %>%
  mutate(
    sum_n = n.x + n.y,
    prob = ((prob.x * (n1/(n1+n2))) + (prob.y * (n2/(n1+n2))))
  ) %>%
  select(value, prob)


#define the sample profile created by two sample mutations
sample_profile <- mut_df %>%
  select(prob)

sample_profile <- as.matrix(sample_profile)
sample_profile


#testing the lm model
sig <- cosmic.signatures[,c("SBS7a","SBS7c")]
sig.labels <- c("sigA", "sigB")
colnames(sig) <- sig.labels

data<- cbind(sample_profile,sig)
data<-as_tibble(data)
model<- lm(prob ~ 0 + sigA + sigB, data=data)
summary(model)

tidy <- model %>% tidy()
(coefs <- tidy$estimate)
```

```{r}
#plot the sample mutational profile
rownames(mut_df) <- mut_df$value
mut_df <- mut_df %>%
  select(prob)
plot_96_profile(mut_df)
```

## Function for generating the sample mutation profiles
Load the Mutation signatures order
```{r}
mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv"
mutational.signature <-  mut.selected
mut_mat <-  read_csv(mut.order)

mut.join <-  mutational.signature %>%
  inner_join(mut_mat)
order.tri <-  unique(mut.join$ext.context)
mut_order <- as_tibble(order.tri)
```


COSMIC Version 2 (*Use the new function below*)

```{r, message=FALSE, warning=FALSE}
#A, B = Signatures.A, Signatures.B
#n1 = the number of mutations that we want to generate from the Signature.A
#n2 = the number of mutations that we want to generate from the Signature.B
generate_sample_prof_1 <- function(A, B, n1, n2){
  cosmic.signatures <- load_cosmic_matrix()
  sig <- cosmic.signatures[,c(A, B)]
  sig.labels <-  c("A","B")
  colnames(sig) <- sig.labels
  
  #generate random mutation samples of Signature 10
  a <- sample(rownames(sig), size = n1, replace = TRUE, prob = as.vector(sig[,"A"]))
  df_a <- as_tibble(a)
  df_a <- df_a %>%
    count(value) %>%
    mutate(
      value = factor(value, levels = order.tri),
      prob = n/sum(n),
    ) %>%
    merge(mut_order, all = TRUE)
  
  df_a[is.na(df_a)] <- 0     # Replace NA with 0
  
  #generate random mutation samples of Signature 6
  b <- sample(rownames(sig), size = n2, replace = TRUE, prob = as.vector(sig[,"B"]))
  df_b <- as_tibble(b)
  df_b <- df_b %>%
    count(value) %>%
    mutate(
      value = factor(value, levels = order.tri),
      prob = n/sum(n)
    ) %>%
    merge(mut_order, all = TRUE)
  
  df_b[is.na(df_b)] <- 0     # Replace NA with 0
  
  #join those two sample mutations
  mut_df <- df_a %>%
    inner_join(df_b, by = "value") %>%
    mutate(
      prob = ((prob.x * (n1/(n1+n2))) + (prob.y * (n2/(n1+n2))))
    ) %>%
    select(value, prob)
  
  #define the sample profile created by two sample mutations
  sample_profile <- mut_df %>%
    select(prob)
  sample_profile <- as.matrix(sample_profile)
  
  sample_profile
}
```


COSMIC Version 3 (new version) with 2 signatures
This function is generating the sample mutational profile based on the selected signatures `A` and `B` and the number of mutations `n1` and `n2` of each signature.`n1` is the number of mutations that we want to generate from the Signature.A `n2` is the number of mutations that we want to generate from the Signature.B

```{r, message=FALSE, warning=FALSE}
#A, B = Signatures.A, Signatures.B
#n1 = the number of mutations that we want to generate from the Signature.A
#n2 = the number of mutations that we want to generate from the Signature.B

generate_sample_prof_2 <- function(A, B, n1, n2){

  cosmic.signatures <- get_known_signatures()
  sig <- cosmic.signatures[,c(A,B)]
  
  #load the mutation order
  mut.selected <- sample.info %>%
    select (sample) %>%
    inner_join (signatures)
  
  mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv"
  mutational.signature <-  mut.selected
  mut_mat <-  read_csv(mut.order)
  
  mut.join <-  mutational.signature %>%
    inner_join(mut_mat)
  order.tri <-  unique(mut.join$ext.context)
  
  rownames(sig) <- order.tri
  mut_order <- as_tibble(order.tri)
  
  
  #generate random mutation samples of Signature A
  a <- sample(rownames(sig), size = n1, replace = TRUE, prob = as.vector(sig[,A]))
  df_a <- as_tibble(a)
  df_a <- df_a %>%
    count(value) %>%
    mutate(
      value = factor(value, levels = order.tri),
      prob = n/sum(n),
    ) %>%
    merge(mut_order, all = TRUE)
  
  df_a[is.na(df_a)] <- 0     # Replace NA with 0
  
  #generate random mutation samples of Signature B
  b <- sample(rownames(sig), size = n2, replace = TRUE, prob = as.vector(sig[,B]))
  df_b <- as_tibble(b)
  df_b <- df_b %>%
    count(value) %>%
    mutate(
      value = factor(value, levels = order.tri),
      prob = n/sum(n)
    ) %>%
    merge(mut_order, all = TRUE)
  
  df_b[is.na(df_b)] <- 0     # Replace NA with 0
  
  #join those two sample mutations
  mut_df <- df_a %>%
    inner_join(df_b, by = "value") %>%
    mutate(
      sum_n = n.x + n.y,
      prob = ((prob.x * (n1/(n1+n2))) + (prob.y * (n2/(n1+n2))))
    ) %>%
    select(value, prob)
  
  
  #define the sample profile created by two sample mutations
  sample_profile <- mut_df %>%
    select(prob)
  
  sample_profile <- as.matrix(sample_profile)
  sample_profile
}
```


COSMIC Version 3 (new version) with multiple signatures

old
```{r, message=FALSE, warning=FALSE}
#sigs = the array of the signatures that we want to put 
#nums = the number of mutations that we want to generate for each signature
generate_sample_prof <- function(sigs, nums){
  #sigs <- c("SBS10a","SBS10b","SBS10c","SBS6")
  #nums <- c(100,100,100,700)
  
  cosmic.signatures <- get_known_signatures()
  sig <- cosmic.signatures[,sigs]
  
  #load the mutation order
  mut.selected <- sample.info %>%
    select (sample) %>%
    inner_join (signatures)
  
  mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv"
  mutational.signature <-  mut.selected
  mut_mat <-  read_csv(mut.order)
  
  mut.join <-  mutational.signature %>%
    inner_join(mut_mat)
  order.tri <-  unique(mut.join$ext.context)
  
  rownames(sig) <- order.tri
  mut_order <- as_tibble(order.tri)
  
  
  df_a <- c()
  #generate from the first signature to the last signatures and merge with the first signature
  for (i in 1:length(sigs)){
    b <- sample(rownames(sig), size = nums[i], replace = TRUE, prob = sig[,i])
    df_b <- as_tibble(b)
    df_b <- df_b %>%
      count(value) %>%
      mutate(
        value = factor(value, levels = order.tri),
        prob = n/sum(n),
      ) %>%
      merge(mut_order, all = TRUE)
    
    df_b[is.na(df_b)] <- 0     # Replace NA with 0
    df_b <- df_b %>%
      select(-value)
    
    rownames(df_b) <- order.tri
    df_b <- as.matrix(df_b)
    
    df_a <- cbind(df_a, df_b)
  }
  
  #multiply with the assigned proportion of the number 
  prob <- c()
  for(i in seq(2,length(sigs)*2,2)){
    p = df_a[,i]*(nums[i/2]/sum(nums))
    prob <- cbind(prob, p)
  }
  
  prob <- rowSums(prob)
  sample_profile <- as.matrix(prob)
  
  colnames(sample_profile) <- c("prob")
  sample_profile
}
```

new
```{r, message=FALSE, warning=FALSE}
#sigs = the array of the signatures that we want to put 
#nums = the number of mutations that we want to generate for each signature

generate_sample_prof <- function(sigs, nums){

  #sigs <- c("SBS10a","SBS10b","SBS10c","SBS6")
  #nums <- c(100,100,100,700)
  
  cosmic.signatures <- get_known_signatures()
  sig <- cosmic.signatures[,sigs]
  
  #load the mutation order
  mut.selected <- sample.info %>%
    select (sample) %>%
    inner_join (signatures)
  
  mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv"
  mutational.signature <-  mut.selected
  mut_mat <-  read_csv(mut.order)
  
  mut.join <-  mutational.signature %>%
    inner_join(mut_mat)
  order.tri <-  unique(mut.join$ext.context)
  
  rownames(sig) <- order.tri
  mut_order <- as_tibble(order.tri)
  
  
  df_a <- c()
  #generate from the first signature to the last signatures and merge with the first signature
  for (i in 1:length(sigs)){
    b <- sample(rownames(sig), size = nums[i], replace = TRUE, prob = sig[,i])
    df_b <- as_tibble(b)
    df_b <- df_b %>%
      count(value) %>%
      mutate(
        value = factor(value, levels = order.tri),
        prob = n/sum(n),
      ) %>%
      merge(mut_order, all = TRUE)
    
    df_b[is.na(df_b)] <- 0     # Replace NA with 0
    df_b <- df_b %>%
      select(-value)
    
    rownames(df_b) <- order.tri
    df_b <- as.matrix(df_b)
    
    df_a <- cbind(df_a, df_b)
  }
  
  #multiply with the assigned proportion of the number 
  prob <- c()
  mut <- c()
  
  for(i in seq(2,length(sigs)*2,2)){
    p = df_a[,i]*(nums[i/2]/sum(nums))
    prob <- cbind(prob, p)
    
    n = df_a[,i-1]*(nums[i/2]/sum(nums))
    mut <- cbind(mut, n)
  }
  
  prob <- rowSums(prob)
  mut <- rowSums(mut)
  sample_profile <- as.matrix(cbind(prob, mut))
  
  sample_profile
}
```

new2
```{r, message=FALSE, warning=FALSE}
#sigs = the array of the signatures that we want to put 
#nums = the number of mutations that we want to generate for each signature

generate_sample_prof <- function(sigs, nums){

  #sigs <- c("SBS10a","SBS10b","SBS10c","SBS6")
  #nums <- c(100,100,100,700)
  
  cosmic.signatures <- get_known_signatures()
  sig <- cosmic.signatures[,sigs]
  
  #load the mutation order
  mut.selected <- sample.info %>%
    select (sample) %>%
    inner_join (signatures)
  
  mut.order="mutational_sig_oec-main/supplied_data/mut_sig.order.csv"
  mutational.signature <-  mut.selected
  mut_mat <-  read_csv(mut.order)
  
  mut.join <-  mutational.signature %>%
    inner_join(mut_mat)
  order.tri <-  unique(mut.join$ext.context)
  
  rownames(sig) <- order.tri
  mut_order <- as_tibble(order.tri)
  
  
  df_a <- c()
  #generate from the first signature to the last signatures and merge with the first signature
  for (i in 1:length(sigs)){
    b <- sample(rownames(sig), size = nums[i], replace = TRUE, prob = sig[,i])
    df_b <- as_tibble(b)
    df_b <- df_b %>%
      count(value) %>%
      mutate(
        value = factor(value, levels = order.tri),
      ) %>%
      merge(mut_order, all = TRUE)
    
    df_b[is.na(df_b)] <- 0     # Replace NA with 0
    df_b <- df_b %>%
      select(-value)
    
    #rownames(df_b) <- order.tri
    df_b <- as.matrix(df_b)
    
    df_a <- cbind(df_a, df_b)
  }
  
  df_a <- rowSums(df_a)
  df_a <- as_tibble(df_a)
  colnames(df_a) <- c("mut")

  sample_profile <- df_a %>%
    mutate(
      prob = mut/sum(mut)
    )
  sample_profile <- as.matrix(sample_profile)
  rownames(sample_profile) <- order.tri
  
  sample_profile
}
```

This function is plotting the sample mutational profile based on the selected signatures and the number of mutations that we want to generate for each signature.
```{r, message=FALSE, warning=FALSE}
#sigs = the array of the signatures that we want to put 
#nums = the number of mutations that we want to generate for each signature

plot_sample_prof <- function(sample_profile){
  
  sample_profile <- sample_profile[,2]
  sample_profile <- as.matrix(sample_profile)
  colnames(sample_profile) <- c("prob")
  
  #plot the matrix or tibble
  plot_96_profile(sample_profile)
}
```

## Linear Regression function
Linear regression function for the COSMIC Version 2 (*Use the new function below*)
```{r}
lm_test_1 <- function(sample_profile, sigA, sigB){
  
  cosmic.signatures <- load_cosmic_matrix()
  sig <- cosmic.signatures[,c(sigA,sigB)]
  cosmic.signatures <- get_known_signatures()
  sig.labels <- c("sigA", "sigB")
  colnames(sig) <- sig.labels
 
  data<- cbind(sample_profile,sig)
  data<-as_tibble(data)
  model<- lm(prob ~ 0 + sigA + sigB, data=data)
  
  tidy <- model %>% tidy()
  tidy$estimate
}
```

COSMIC Version 3
This Function generate the simple Linear regression model for the generated sample mutational profile based on two signatures Signature.A and Signature.B.
```{r}
lm_test_2 <- function(sample_profile, sigA, sigB){
  
  cosmic.signatures <- get_known_signatures()
  sig <- cosmic.signatures[,c(sigA,sigB)]
  cosmic.signatures <- get_known_signatures()
  sig.labels <- c("sigA", "sigB")
  colnames(sig) <- sig.labels
 
  data<- cbind(sample_profile,sig)
  data<-as_tibble(data)
  model<- lm(prob ~ 0 + sigA + sigB, data=data)
  
  tidy <- model %>% tidy()
  tidy$estimate
}
```


COSMIC Version 3
This Function generate the Linear regression model with positive constraints for the generated sample mutational profile based on the selected multiple signatures.
```{r}
lm_test <- function(sample_profile, sigs, constrs){
  
  cosmic.signatures <- get_known_signatures()
  sig <- cosmic.signatures[,sigs]
  
  df <- cbind(sample_profile,sig)
  df <- as_tibble(df)
  
  constraints <- str_c(constrs[1], ' < ', sigs[1], ', ', sigs[1], ' < ', constrs[2])
  for (i in 2:length(sigs)){
    constraint <- str_c(constrs[2*i-1], ' < ', sigs[i], ', ', sigs[i], ' < ', constrs[2*i])
    constraints <- str_c(constraints, ', ', constraint)
  }
  
  model<- ConsReg(formula = prob ~ 0 + ., data = df,
            family = 'gaussian',
            constraints = constraints
            )
  
  model$coefficients
  #coeff <- as.tibble(model$coefficients)
  #coeff$value
}
```

new
```{r}
estimate_coeff <- function(sample_profile, sigs, tissue_type, scale, quantiles, mb_size){
  
  #parameters testing
  #sample_profile <- generate_sample_prof(c("SBS1","SBS10b","SBS6"),c(100,300,600))
  total_mut <- unname(colSums(sample_profile)[1])
  
  sample_profile <- as.matrix(sample_profile[,2])
  colnames(sample_profile) <- c("prob")
  
  #sigs <- c("SBS1","SBS10b","SBS6")
  #tissue_type <- "BRCA"
  #scale <- "exome"
  #quantiles <- c("5%","95%")
  #mb_size <- 50
  
  #pulling constraints
  constraints_ds <- read_csv("Data/constraints_ds.csv")
  constr1 <- constraints_ds %>%
    filter(Tissue == tissue_type, Signature %in% sigs, Scale == scale, Quantile %in% quantiles) 
  constrs <- c(constr1$Value)
  
  #making dataset for the model
  cosmic.signatures <- get_known_signatures()
  sig <- cosmic.signatures[,sigs]
  
  df <- cbind(sample_profile,sig)
  df <- as_tibble(df)
  
  constraints <- str_c(constrs[1]*(mb_size/total_mut), ' < ', sigs[1], ', ', sigs[1], ' < ', constrs[2]*(mb_size/total_mut))
  for (i in 2:length(sigs)){
    constraint <- str_c(constrs[2*i-1]*(mb_size/total_mut), ' < ', sigs[i], ', ', sigs[i], ' < ', constrs[2*i]*(mb_size/total_mut))
    constraints <- str_c(constraints, ', ', constraint)
  }
  
  model<- ConsReg(formula = prob ~ 0 + ., data = df,
            family = 'gaussian',
            constraints = constraints
            )
  
  model$coefficients
  #coeff <- as.tibble(model$coefficients)
  #coeff$value
}
```

Function testing code
```{r}
#A, B = Signatures.A, Signatures.B
#n1 = the number of mutations that we want to generate from the Signature.A
#n2 = the number of mutations that we want to generate from the Signature.B

#function for the old COSMIC version 
sample_profile <- generate_sample_prof_1(10,6,4000,6000)
lm_test_1(sample_profile,10,6)

sample_profile1 <- generate_sample_prof_2("SBS10a","SBS6", 2000, 8000)
plot_sample_prof(sample_profile1)
lm_test(sample_profile1,c("SBS10a","SBS6"), c(0.1, 1, 0.1, 1))

#2 sigs
sample_profile <- generate_sample_prof(c("SBS10a","SBS6"),c(3000,7000))
plot_sample_prof(sample_profile)
estimate_coeff(sample_profile, sigs = c("SBS10a","SBS6"), tissue_type = "BRCA", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#3 sigs
sample_profile <- generate_sample_prof(c("SBS1","SBS10b","SBS6"),c(200,300,500))
plot_sample_prof(sample_profile)
estimate_coeff(sample_profile, sigs = c("SBS1","SBS10b","SBS6"), tissue_type = "BRCA", scale = "exome", quantiles = c("40%", "50%"), mb_size = 50)


brca <- constraints_ds %>%
  filter(Tissue == "BRCA") 
unique(brca$Signature)

#4 sigs
sample_profile <- generate_sample_prof(c("SBS10a","SBS10b","SBS1","SBS6"),c(300,100,200,400))
plot_sample_prof(sample_profile)
estimate_coeff(sample_profile, sigs = c("SBS10a","SBS10b","SBS1","SBS6"), tissue_type = "BRCA", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)

#5 sigs
sample_profile <- generate_sample_prof(c("SBS10a","SBS10b","SBS1","SBS6", "SBS2"),c(200,300,100,300,100))
plot_sample_prof(sample_profile)
estimate_coeff(sample_profile, sigs = c("SBS10a","SBS10b","SBS1","SBS6", "SBS2"), tissue_type = "BRCA", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)
```


These are the signatures that the constraints are unknown in the COSMIC website.
If we want to include the signatures below, we might need to use the function that the constraints can be set manually.
Also the old version of generate_sample_prof needs to be used here.
```{r, eval=FALSE}
unknown <- c("SBS10c", "SBS10d", "SBS25", "SBS48", "SBS84", "SBS85", "SBS86", "SBS87", "SBS88", "SBS89", "SBS90", "SBS91", "SBS92", "SBS93", "SBS94")
```
```{r, eval=FALSE}
sample_profile <- generate_sample_prof(c("SBS10c","SBS10d","SBS1","SBS6"),c(300,100,200,400))
plot_sample_prof(sample_profile)
#estimate_coeff(sample_profile, sigs = c("SBS10c","SBS10d","SBS1","SBS6"), tissue_type = "BRCA", scale = "exome", quantiles = c("5%", "95%"), mb_size = 50)
lm_test(sample_profile, sigs = c("SBS10c","SBS10d","SBS1","SBS6"), constrs = c(0.1, 1, 0.1, 1, 0.1, 1, 0.1, 1))
```

